% -*- root: main.tex -*-



\chapter{Power operations}

Our goal in this Appendix is to give a tour of the interaction of the $\sigma$--orientation with a topic of modern research, the theory of $E_\infty$ ring spectra, in a manner consistent with the rest of the topics in this book.  Because the theory of $E_\infty$ ring spectra (in particular: their algebraic geometry) is still very much developing, we have no hope of stating results in their maximum strength or giving a completely clear picture---as of this writing, the maximum strength is unknown and the picture is still resolving.  Although $E_\infty$ ring spectra themselves were introduced decades ago, we will even avoid giving a proper definition of them here, instead referring to the original work of May and collaborators~\cite{EKMM} and the more recent work of Lurie~\cite[Chapter 7]{LurieHA} for a proper treatment.  In acknowledgement of this underwhelming level of rigor, we have downgraded our discussion from a Case Study to an Appendix.

As far as we are concerned, $E_\infty$ ring spectra arise in order to solve the following problem: given two ring spectra $R$ and $S$ in the homotopy category, the set of homotopy classes of ring maps $\CatOf{RingSpectra}(R, S)$ forms a subset of the set of all homotopy classes $[R, S] = \pi_0 \CatOf{Spectra}(R, S)$, selected by a homomorphism condition.  There is no meaningful way to enrich this to a \emph{space} of ring spectrum maps from $R$ to $S$, which inhibits us from understanding an obstruction theory for ring spectra, i.e., approximating $R$ by ``nearby'' ring spectra $R'$ in a way that relates $\CatOf{RingSpectra}(R, S)$ and $\CatOf{RingSpectra}(R', S)$ by a fiber sequence.

The extra data that accomplishes this mapping space feat turns out to be an explicit naming of the homotopies controlling the associativity and commutativity of the ring spectrum multiplication, which are subject to highly intricate compatibility conditions.\footnote{This is rather analogous to the extra data required on a \emph{space}, beyond just a multiplication, which allows one to use the bar construction to assemble a delooping.}\footnote{The high degree of intricacy accomplishes this goal of constructing mapping spaces, but it interacts strangely with the classical notion of a ring spectrum in the homotopy category: there are ring spectrum maps that admit \emph{no} enrichment to an $E_\infty$ map, and there are ring spectrum maps that admit \emph{multiple} enrichments to $E_\infty$ maps.}  Again, rather than spell this out, it suffices for our purposes to say that there is such a notion of a structured ring spectrum that begets a mapping space between two such.  Additionally, we record the following omnibus theorem as indication that this program overlaps with the one we have been describing already:
\begin{theorem}
The following are examples of $E_\infty$ ring spectra:
\begin{itemize}
    \item (\cite[Section VIII.1]{MayRingSpacesSpectra}) The classical $K$--theories $KU$ and $KO$.
    \item (\cite[Section VIII.1]{MayRingSpacesSpectra}) The Eilenberg--Mac Lane spectra $HR$.
    \item (\cite[Corollary 7.6--7]{GoerssHopkins}) The Morava $E$--theories $E_\Gamma$ and their fixed point spectra.\footnote{Notably, the Morava $K$--theories are \emph{not} $E_\infty$ rings at finite heights, in view of \Cref{HinftyRingsModp}.}
    \item (\cite[Section IV.3]{MayRingSpacesSpectra}) The Thom spectra arising from the $J$--homomorphism, including $MO$, $MSO$, $M\Spin$, $M\String$, $MU$, $MSU$, and $MU[6, \infty)$.
    \item (\citeme{Behrens}, cf.\ \Cref{ConstructionOfTMFSection}) The spectra $\TMF$, $\Tmf$, and $\tmf$. \qed
\end{itemize}
\end{theorem}

The forgetful map from $E_\infty$ rings down to ring spectra in the homotopy category factors through an intermediate category, that of $H_\infty$ ring spectra, which captures the extra factorizations expressing these associativity and commutativity relations.  Specifically, recall the following definition from the discussion in \Cref{QuillenPowerOpnsSection}:

\begin{definition}[{\cite[Definition I.3.1]{BMMS}, cf.\ \Cref{QuillenPowerOpnsSection}}]
An $H_\infty$ ring spectrum is a ring spectrum $E$ equipped with factorizations $\mu_n$ as in
\begin{center}
\begin{tikzcd}
E^{\sm n} \arrow["\mu"]{r} \arrow{d} & E \\
E^{\sm n}_{h\Sigma_n} \arrow{ru}[description]{\mu_n},
\end{tikzcd}
\end{center}
which are subject to compatibilities induced by the inclusions $\Sigma_n \times \Sigma_m \subseteq \Sigma_{n+m}$ and the inclusions $\Sigma_n \wr \Sigma_m \subseteq \Sigma_{nm}$.
\end{definition}

\begin{lemma}
Each $E_\infty$ ring spectrum gives rise to an $H_\infty$ ring spectrum in the homotopy category. \qed
\end{lemma}

\noindent We care about this secondary definition because our results thus far have all concerned the cohomology of spaces, which is, at its core, a calculation at the level of \emph{homotopy classes}.  This is therefore as much of the $E_\infty$ structure as one could hope would interact with our analyses in the preceding Case Studies.

In \Cref{QuillenPowerOpnsSection}, \Cref{StabilizingTheMUSteenrodOps}, and \Cref{CalculationOfMUStarSection}, we introduced an $H_\infty$ ring structure on $MU$ and used it to make a calculation of the coefficient ring $MU_*$.  Our primary goal in this Appendix is to introduce an $H_\infty$ ring structure on certain chromatically interesting spectra, including Morava's theories $E_\Gamma$, and to describe the compatibility laws arising from intertwining these two $H_\infty$ structures.  The culminating result is as follows:

\begin{theorem}[{cf.\ \Cref{AHSHinftyResultForEthy}}]
An orientation $MU[6, \infty) \to E_\Gamma$ is $H_\infty$ if and only if the induced cubical structure is ``norm-coherent'' (cf.\ \Cref{NormCoherentDefn}). \qed
\end{theorem}

\noindent Before addressing this, we discuss in \Cref{CharacterTheorySection} an important phenomenon: after deleting certain forms of torsion, the Morava $E_\Gamma$--homology of a finite spectrum can be well--approximated by its Morava $E_{\Gamma'}$--homology where $\Gamma'$ satisfies $\height{\Gamma'} < \height{\Gamma}$.  This is interesting in its own right, and we will quickly see that the precise form of the approximation bears directly on the study of power operations.  Finally, with the homotopy category exposed, we give a summary of the known results about $E_\infty$ orientations themselves in \Cref{ConstructionOfTMFSection}, where we summarize the construction of a spectrum $\TMF$, and in \Cref{JuvitopTalkSection}, where we summarize what goes into giving it a $\String$--orientation.













\section{Rational phenomena: character theory for Lubin--Tate spectra}\label{CharacterTheorySection}

\todo[inline]{Nat is going to revamp arXiv:1308.1414 for use here.}

There's a sufficient amount of reliance on character theory in Matt's thesis that we should talk about it.  You should write that action and then backtrack here to see what you need for it.

See Morava's \textit{Local fields} paper

\begin{remark}
Theorem 2.6 of Greenlees--Strickland for a nice transchromatic perspective.  See also work of Stapleton and Schlank--Stapleton, of course.\todo{Flesh this out.}
\end{remark}


------

\begin{theorem}\citeme{Theorem A}
Let $E$ be any complex-oriented cohomology theory.  Take $G$ to be a finite group and let $\CatOf{Ab}_G$ be the full subcategory of the orbit category of $G$ built out of abelian subgroups of $G$.  Finally, let $X$ be a finite $G$--CW complex.  Then, each of the natural maps \[E^*(EG \times_G X) \to \lim_{A \in \CatOf{Ab}_G} E^*(EG \times_A X) \to \int_{A \in \CatOf{Ab}_G} E^*(BA \times X^A)\] becomes an isomorphism after inverting the order of $G$.  In particular, there is an isomorphism
\[\pushQED{\qed}
\frac{1}{|G|} E^* BG \to \lim_{A \in \CatOf{Ab}_G} \frac{1}{|G|} E^* BA. \qedhere
\popQED\]
\end{theorem}

This is an analogue of Artin's theorem:
\begin{theorem}
There is an isomorphism
\[\pushQED{\qed}
\frac{1}{|G|} R(G) \to \lim_{C \in \CatOf{Cyclic}_G} \frac{1}{|G|} R(C). \qedhere
\popQED\]
\end{theorem}


------

HKR intro material connecting Theorem A to character theory:

Recall that classical characters for finite groups are defined in the following situation: take $L = \Q^{\mathrm{ab}}$ to be the smallest characteristic $0$ field containing all roots of unity, and for a finite group $G$ let $Cl(G; L)$ be the ring of class functions on $G$ with values in $L$.  The units in the profinite integers $\widehat{\Z}$ act on $L$ as the Galois group over $\Q$, and since $G = \CatOf{Groups}(\widehat{\Z}, G)$ they also act naturally on $G$.  Together, this gives a conjugation action on $Cl(G; L)$: for $\phi \in \widehat{\Z}$, $g \in G$, and $\chi \in Cl(G; L)$, one sets \[(\phi \cdot \chi)(g) = \phi(\chi(\phi^{-1}(g))).\]  The character map is a ring homomorphism \[\chi: R(G) \to Cl(G; L)^{\widehat{\Z}},\] and this induces isomorphisms \[\chi: L \otimes R(G) \xrightarrow{\simeq} Cl(G; L)\] and even \[\chi: \Q \otimes R(G) \xrightarrow{\simeq} Cl(G; L)^{\widehat{\Z}}.\]

Now take $E = E_\Gamma$ to be a Morava $E$--theory of finite height $d = \height(\Gamma)$.  Take $E^*(B\Z_p^d)$ to be topologized by $B(\Z/p^j)^d$.  A character $\alpha: \Z_p^d \to S^1$ will induce a map $\alpha^*: E^* \CP^\infty \to E^* B\Z_p^d$.  We define $L(E^*) = S^{-1} E^*(B\Z_p^d)$, where $S$ is the set of images of a coordinate on $\CP^\infty_E$ under $\alpha^*$ for nonzero characters $\alpha$.  Note that this ring inherits an $\operatorname{Aut}(\Z_p^d)$ action by $E^*$--algebra maps.

The analogue of $Cl(G; L)$ will be $Cl_{d,p}(G; L(E^*))$, defined to be the ring of functions $\chi: G_{d, p} \to L(E^*)$ stable under $G$--orbits.  Noting that \[G_{d,p} = \operatorname{Hom}(\Z_p^d, G),\] one sees that $\operatorname{Aut}(\Z_p^d)$ acts on $G_{d,p}$ and thus on $Cl_{d,p}(G; L(E^*))$ as a ring of $E^*$--algebra maps: given $\phi \in \operatorname{Aut}(\Z_p^d)$, $\alpha \in G_{d,p}$, and $\chi \in Cl_{d,p}(G; L(E^*))$ one lets \[(\phi \cdot \chi)(\alpha) = \phi(\chi(\phi^{-1}(\alpha))).\]

Now we introduce a finite $G$--CW complex $X$.  Let \[\operatorname{Fix}_{d, p}(G, X) = \coprod_{\alpha \in \operatorname{Hom}(\Z_p^d, G)} X^{\operatorname{im} \alpha}.\]  This space has commuting actions of $G$ and $\operatorname{Aut}(\Z_p^d)$.  We set \[Cl_{d, p}(G, X; L(E^*)) = L(E^*) \otimes_{E^*} E^*(\operatorname{Fix}_{d,p}(G, X))^G,\] which is again an $E^*$--algebra acted on by $\operatorname{Aut}(Z_p^d)$.  We define the character map ``componentwise'': a homomorphism $\alpha \in \operatorname{Hom}(\Z_p^d, G)$ induces \[E^*(EG \times_G X) \to E^*(B\Z_p^d) \otimes_{E^*} E^*(X^{\operatorname{im} \alpha}) \to L(E^*) \otimes_{E^*} E^*(X^{\operatorname{im} \alpha}).\]  Taking the direct sum over $\alpha$, this assembles into a map \[\chi_{d,p}^G: E^*(EG \times_G X) \to Cl_{d,p}(G, X; L(E^*))^{\operatorname{Aut}(Z_p^d)}.\]
\todo{Nat taught you how to say all these things with $p$--adic tori, which was \emph{much} clearer.}
\begin{theorem}\citeme{Theorem C}
The invariant ring is $L(E^*)^{\operatorname{Aut}(\Z_p^d)} = p^{-1} E^*$, and $L(E^*)$ is faithfully flat over $p^{-1} E^*$.\todo{Checking this invariant ring claim is easiest done by comparing the functors the two things corepresent.}  The character map $\chi_{d,p}^G$ induces isomorphisms
\begin{align*}
\chi_{d,p}^G \co L(E^*) \otimes_{E^*} E^*(EG \times_G X) & \xrightarrow{\simeq} Cl_{d,p}(G, X; L(E^*)), \\
\chi_{d,p}^G \co p^{-1} E^*(EG \times_G X) & \xrightarrow{\simeq} Cl_{d,p}(G, X; L(E^*))^{\operatorname{Aut}(\Z_p^d)}.
\end{align*}
In particular, when $X = *$, there are isomorphisms
\begin{align*}
\chi_{d,p}^G \co L(E^*) \otimes_{E^*} E^*(BG) & \xrightarrow{\simeq} Cl_{d,p}(G; L(E^*)), \\
\chi_{d,p}^G \co p^{-1} E^*(BG) & \xrightarrow{\simeq} Cl_{d,p}(G; L(E^*))^{\operatorname{Aut}(\Z_p^d)}. \qed
\end{align*}
\end{theorem}

------

Jack gives an interpretation of this in terms of formal $\sheaf{O}_L$--modules.

------

I also have this summary of Nat's of the classical case:

It's not easy to decipher if you weren't there for the conversation, but here's my take on it. First, the map we wrote down today was the non-equivariant chern character: it mapped non-equivariant $KU \otimes \Q$ to non-equivariant $H\Q$, periodified. The first line on Nat's board points out that if you use this map on Borel-equivariant cohomology, you get nothing interesting: $K^0(BG)$ is interesting, but $H\Q^*(BG) = H\Q^*(*)$ collapses for finite $G$. So, you have to do something more impressive than just directly marry these two constructions to get something interesting.

That bottom row is Nat's suggestion of what ``more interesting'' could mean. (Not really his, of course, but I don't know who did this first. Chern, I suppose.) For an integer $n$, there's an evaluation map of (forgive me) topological stacks \[* \mmod (\Z/n) \times \operatorname{Hom}(* \mmod (\Z/n), * \mmod G) \xrightarrow{\mathrm{ev}} * \mmod G\] which upon applying a global-equivariant theory like $K_G$ gives \[K_{\Z/n}(*) \otimes K_G(\coprod_{\text{conjugacy classes of $g$ in $G$}} *) \xleftarrow{ev^*} K_G(*).\]

Now, apply the genuine $G$-equivariant Chern character to the $K_G$ factor to get \[K_{\Z/n}(*) \otimes H\Q_G(\coprod *) \from K_{\Z/n}(*) \otimes K_G(\coprod *),\] where the coproduct is again taken over conjugacy classes in G. Now, compute $K_{\Z/n}(*) = R(\Z/n) = \Z[x] / (x^n - 1)$, and insert this calculation to get \[K_{\Z/n}(*) \otimes H\Q_G(\coprod *) = \Q(\zeta_n) \otimes (\bigoplus_{\text{conjugacy classes}} \Q),\] where $\zeta_n$ is an $n${\th} root of unity.  As $n$ grows large, this selects sort of the part of the complex numbers $\C$ that the character theory of finite groups cares about, and so following all the composites we've built a map \[K_G(*) \to \C \otimes (\bigoplus_{\text{conjugacy classes}} \C).\]  The claim, finally, is that this map sends a $G$-representation (thought of as a point in $K_G(*)$) to its class function decomposition.














\section{Orientations and power operations}\label{PowerOpnsSection}

Our introduction of $E_\infty$ rings also automatically introduces a few interesting accompanying functors:
\begin{center}
\begin{tikzcd}
\CatOf{Spaces} \arrow["E^{(-)_+}", bend left]{rr} & \CatOf{Modules}_E \arrow["\P_E", shift left=0.4em]{r} \arrow[shift left=0.4em]{d} & \EinftyRings_E \arrow[shift left=0.4em]{l} \arrow[shift left=0.4em]{d} \\
& \CatOf{Spectra} \arrow[shift left=0.4em, "(-) \sm E"]{u} \arrow["\P", shift left=0.4em]{r} & \EinftyRings. \arrow[shift left=0.4em, "(-) \sm E"]{u} \arrow[shift left=0.4em]{l}
\end{tikzcd}
\end{center}
The first functor sends a space $X$ to its spectrum of $E$--cochains $E^{\Susp^\infty_+ X}$, and the other two functors form a free/forgetful monad resolving a mapping space in $\EinftyRings_E$ by a sequence of mapping spaces in $\CatOf{Modules}_E$.  These kinds of functors are familiar to us from the discussion of contexts in \Cref{StableContextLecture} and \Cref{UnstableContextsSection}, and the recipe applied in those situations gives an analogous story here.  First, there is a natural map \[\CatOf{Spaces}(*, X) \to \EinftyRings_E(E^{X_+}, E),\] which one hopes is an equivalence under (often very strong) hypotheses on $E$ and on $X$.\footnote{It is an unpublished theorem of Hopkins and Lurie that if $X$ is a space admitting a finite Postnikov system with at most $\height \Gamma$ stages and involving only finite groups, then the natural map $F(*, X) \to \EinftyRings_{E_\Gamma/}(E_\Gamma^{X_+}, E_\Gamma)$ is an equivalence.}  Second, the adjunction gives a mechanism for resolving $E^{X_+}$, which feeds into a spectral sequence computing this right-hand mapping space.  The functors $\P$ and $\P_E$ can be given by explicit formulas:
\begin{align*}
\P(X) & = \bigvee_{j=0}^\infty X^{\sm j}_{h\Sigma_j}, &
\P_E(M) & = \bigvee_{j=0}^\infty M^{\sm_E j}_{h\Sigma_j}.
\end{align*}
Finally, we expect the homotopy groups of this resolution to form a quasicoherent sheaf over a suitable \emph{$E_\infty$ context}, which arises as the simplicial scheme associated to this resolution in the case where $X$ is a point.  In this case, we can explicitly name some of the terms in this resolution: the bottom two stages take the form
\begin{center}
\begin{tikzcd}
\EinftyRings_E(E, E) \\
\EinftyRings_E(\P_E(E), E) \arrow{u} \arrow{r} & \EinftyRings_E(\P_E^2(E), E) \arrow[shift left=0.4em]{l} \arrow[shift right=0.4em]{l} \arrow[shift left=0.4em] {r} \arrow[shift right=0.4em]{r} & \arrow[shift left=0.8em]{l} \arrow{l} \arrow[shift right=0.8em]{l} \cdots.
\end{tikzcd}
\end{center}
The available adjunctions give a more explicit presentation of these terms:
\begin{align*}
\EinftyRings_E(\P_E(E), E) & \simeq \CatOf{Modules}_E(E, E) \\
& \simeq \CatOf{Spectra}(\S, E),
\intertext{which on homotopy groups computes the coefficient ring of $E$, and}
\EinftyRings_E(\P_E^2(E), E) & \simeq \CatOf{Modules}_E(\P_E(E), E) \\
& \simeq \CatOf{Modules}_E\left(\bigvee_{j=0}^\infty E^{\sm_E j}_{h\Sigma_j}, E\right) \\
& \simeq \CatOf{Spectra}\left(\bigvee_{j=0}^\infty \S^{\sm j}_{h\Sigma_j}, E\right) \\
& \simeq \prod_{j=0}^\infty \CatOf{Spectra}\left(B\Sigma_j, E\right),
\end{align*}
which on homotopy groups is made up of a product of the cohomology rings $E^*(B\Sigma_j)$.  The higher terms track the compositional behavior of these summands.

\begin{remark}[{\cite{BousfieldUnstableLocalization,MahowaldThompson}}]
One of the first places these ideas appear in the literature is in work of Mahowald and Thompson.  Bousfield defined an unstable local homotopy type associated to a (simply-connected) space and a homology theory.  In the case of the space $S^{2n-1}$ and $p$--adic $K$--theory, Mahowald and Thompson calculated that $L_K S^{2n-1}$ appears as the homotopy fiber \[L_K S^{2n-1} \to L_K \Loops^\infty \Sigma^\infty S^{2n-1} \to L_K \Loops^\infty \Sigma^\infty ((S^{2n-1})^{\sm p}_{h\Sigma_p}),\] which is an abbreviated form of the monadic resolution described above.
\end{remark}

\begin{remark}
\citeme{This should be in...Charles's work?}
In general, if $E^* B\Sigma_j$ is sufficiently nice, then the the $E^2$--page of the monadic descent spectral sequence computes the derived functors of derivations, taken in a suitable category of monad-algebras for the monad specifying the behavior of power operations.
\end{remark}






\subsection*{Strickland's theorems}

We thus set out to understand the formal schemes constituting the $E_\infty$ context associated to Morava $E$--theory.  As described in \Cref{UnstableContextsSection} and \Cref{UnstableAlgebraicModelSection}, the correct language for these phenomena are schemes defined on Hopf rings, together with the adjunction between classical rings and Hopf rings consisting of the $\ast$--square--zero extension functor and the $\ast$--indecomposables functor.  The rings $E^0 B\Sigma_j$ assemble into a Hopf ring using the following structure:
\begin{itemize}
    \item The $\ast$--product comes from the stable transfer maps $B\Sigma_{i+j} \to B\Sigma_i \times B\Sigma_j$.
    \item The $\circ$--product comes from the diagonal maps $B\Sigma_j \to B\Sigma_j \times B\Sigma_j$.
    \item The diagonal comes from the block-inclusion maps $B\Sigma_i \times B\Sigma_j \to B\Sigma_{i+j}$.
\end{itemize}

\begin{definition}
Accordingly, we set the \textit{natural $E_\infty$ context} to be \[\Econtext{E_\Gamma} = \SpH E^0 B\Sigma_*.\]
\end{definition}

The effect of this functor on classical rings is given by \Cref{HopfRingsAndRingsAdjunction}: $\Econtext{E_\Gamma}(T) = \CatOf{Algebras}_{E_\Gamma^0}(Q^* E^0 B\Sigma_*, T)$, where \[Q^* E^0 B\Sigma_* = \frac{E^0 B\Sigma_*}{\im\left(\Tr_{\Sigma_{*_1} \times \Sigma_{*_2}}^{\Sigma_{*_1 + *_2}}\co E^0 B\Sigma_{*_1} \times E^0 B\Sigma_{*_2} \to E^0 B\Sigma_{*_1 + *_2}\right)}.\]  The ideal appearing in this equation is called the \textit{transfer ideal}, written $I_{\Tr}$.

\begin{remark}
In terms of the descent spectral sequence described in the previous subsection, all of the $*$--decomposables are in the image of the $d^1$--differential, so already do not contribute to the $E^2$--page of the spectral sequence.
\end{remark}

We dissect this Hopf ring spectrum by considering the $j${\th} graded piece of $\Econtext{E_\Gamma}$ as restricted to classical rings, i.e., by understanding the formal schemes $\Spec (E^0 B\Sigma_j / I_{\Tr})$ for individual indices $j$.

\begin{example}
To gain a foothold, it is helpful to further specialize to a particular case---say, $j = p$.  In light of the results of \Cref{CharacterTheorySection}, we might begin by analyzing the (maximal) abelian subgroups of $\Sigma_p$, of which the only $p$--locally interesting one is the transitive subgroup $C_p \subseteq \Sigma_p$.  In \Cref{KtheoryConvertsTorsionToTorsion}, we calculated $BS^1[p]_E = \CP^\infty_E[p]$, and we now make the further observation that the regular representation map $\rho\co BS^1[p] \to BU(p)$ induces the following map on cohomological formal schemes:
\begin{center}
\begin{tikzcd}
BS^1[p]_E \arrow["\rho_E"]{r} & BU(p)_E \\
\InternalHom{FormalGroups}(\Z/p, \CP^\infty_E) \arrow{r} \arrow[equal]{u} & \Div_p^+ \CP^\infty_E \arrow[equal]{u},
\end{tikzcd}
\end{center}
where the bottom arrow sends such a homomorphism to its image divisor.  This map belongs to a larger diagram of schemes:
\begin{center}
\begin{tikzcd}
\Spf E^0 B\Sigma_p / I_{\Tr} \arrow{r} & (B\Sigma_p)_E \arrow{r} & BU(p)_E \\
\Spf E^0 BC_p / I_{\Tr} \arrow{r} \arrow{u} & (BC_p)_E \arrow{u} \arrow{ru}.
\end{tikzcd}
\end{center}
The effect of killing the transfer ideal in $(BC_p)_E$ is to force the image divisor to be a subdivisor of $\CP^\infty_E[p]$\todo{Explain why.} (i.e., the zero homomorphism is disallowed), and this subscheme of homomorphisms is written $\Level(\Z/p, \CP^\infty_E)$.  Finally, passing to $B\Sigma_p$ from $BC_p$ exactly destroys the choice of generator of $\Z/p$, i.e., it encodes passing from the homomorphism to the image divisor.  This winds up giving an isomorphism \[\Spf E^0 B\Sigma_p / I_{\Tr} \cong \Sub_p \CP^\infty_E,\] where $\Sub_p$ denotes the subscheme of $\Div_p^+$ consisting of those effective Weil divisors of rank $p$ which are subgroup divisors.
\end{example}

The broad features of this example hold for a general index $j$.
\begin{definition}
The \textit{abelian $E_\infty$ context} is formed by considering the inclusions \[\bigvee_{\substack{A \le \Sigma_j \\ \text{$A$ abelian}}} BA \to B\Sigma_j.\]
\end{definition}
\noindent A consequence of \Cref{CharacterTheorySection} is that this map is \emph{injective} on Morava $E$--cohomology, so that we can understand the natural $E_\infty$ context in terms of this larger object.  A benefit to this auxiliary context is that we can already predict its behavior in Morava $E$--theory: the cohomological formal scheme associated to an abelian group can be presented as an internal scheme of group homomorphisms, just as above.  Using this auxiliary context, Strickland has proven the following results:

\begin{theorem}[{\cite[Theorem 1.1]{StricklandEthyOfBSigma}}]
There is an isomorphism \[\Spf E^0 B\Sigma_j / I_{\Tr} \cong \Sub_j \CP^\infty_E,\] where $\Sub_j$ denotes the subscheme of $\Div_j^+$ consisting of those effective Weil divisors of rank $j$ which are subgroup divisors.\footnote{If $j$ is not a power of $p$, this is the terminal scheme.} \qed
\end{theorem}

\begin{theorem}\citeme{Strickland somewhere}
For a finite abelian group $A$, there is an isomorphism \[\Spf E^0 BA / I_{\Tr} \cong \Level(A^*, \CP^\infty_E),\] where $\Level(A^*, \CP^\infty_E)$ denotes the subscheme of $\InternalHom{FormalGroups}(A^*, \CP^\infty_E)$ subject to the condition that $A^*[n]$ forms a subdivisor of $\CP^\infty_E[n]$.\footnote{If $A[p] \cong (\Z/p)^{\times k}$ has $k > \height{\Gamma}$, then this is the terminal scheme.}  \qed
\end{theorem}

\begin{remark}
An important piece of intuition about the schemes $\Level(A^*, \CP^\infty_E)$ is that they form a kind of replacement for the nonexistent ``scheme of monomorphisms''.  Specifically, the $p$--series for a Lubin--Tate universal deformation group is only once $x$--divisible, and hence the divisor $\G[p]$ only contains the divisor $[0]$ with multiplicity one.\footnote{This kind of reasoning applies to domains of characteristic $0$ generally.}  This excludes noninjective morphisms in this case.  On the other hand, the only subgroups of the formal group restricted to the special fiber are of the form $p^m \cdot [0]$.  In particular, any level structure on $\G$ restricts to a morphism with this image divisor at the special fiber, and hence functoriality considerations force us to count these---which are \emph{not} images of monomorphisms---as level structures as well.
\end{remark}

\begin{remark}[{\cite{StricklandFiniteSubgps}}]
The schemes $\Sub_j \G$ and $\Level(A, \G)$ are known to possess many very pleasant algebraic properties: they are finite and free of predictable rank, they have Galois descent properties, the schemes $\Level(A, \G)$ are all reduced, there are important decompositions coming from presenting a subgroup scheme as a flag of smaller subgroups, \ldots.  Indeed, these algebraic results form important ingredients to the proof of the connection with homotopy theory~\cite[Section 9]{StricklandEthyOfBSigma}.
\end{remark}










\subsection*{Isogenies and the Lubin--Tate moduli}\label{IsogeniesSection}

In this subsection, we seek a comparison of the natural $E_\infty$ context and the unstable context considered in \Cref{UnstableContextsSection}.  Our model for the unstable context in \Cref{UnstableAlgebraicModelSection} focuses on the effect of unstable operations on the cohomology of $\CP^\infty$, as summarized in the following result:

\begin{lemma}[{mild extension of \Cref{UnstableCooperationsChapter}}]
There is an isomorphism \[\Spec Q^* \pi_* L_\Gamma(E \sm E) \cong \InternalHom{FormalGroups}(\CP^\infty_E, \CP^\infty_E). \qed\]
\end{lemma}

\noindent In order to form a comparison map between these two contexts, we will want algebraic constructions that trade a subgroup divisor (i.e., a point in the natural $E_\infty$ context) for a formal group endomorphism (i.e., a point in the unstable context).  It will be useful to phrase our ideas in the language of \emph{isogenies}.

\begin{definition}[{\cite[Definition 5.17]{StricklandFSFG}}]
Take $C$ and $D$ to be formal curves over $X$.  A map $f\co C \to C'$ is an \textit{isogeny} (of degree $d$) when the induced map $C \to C \times_X C'$ exhibits $C$ as a divisor (of rank $d$) on $C \times_X C'$ as $C'$--schemes.
\end{definition}

\begin{remark}[{cf.\ \Cref{DivHasPushforwards}}]
In this case, a divisor $D$ on $C'$ gives rise to a divisor $f^* D$ on $C$ by scheme-theoretic pullback:
\begin{center}
\begin{tikzcd}
f^* D \arrow{r} \arrow{d} & D \arrow{d} \\
C \arrow["f"]{r} & C',
\end{tikzcd}
\end{center}
altogether inducing a map \[f^*\co \Div_n^+ C' \to \Div_{nd}^+ C.\]  This map interacts with pushforward by $f_* f^* D = d \cdot D$, where $d$ is the degree of the isogeny.
\end{remark}

The usual source of examples of isogenies are polynomial maps between curves.  In fact, this is close to the general case, and the following result is the source of much intuition:
\begin{lemma}[{Weierstrass preparation, \cite[Section 5.2]{StricklandFSFG}}]
Let $R$ be a complete local ring.  Every degree $d$ isogeny $f\co \A^1_R \to \A^1_R$ admits a unique factorization as a coordinate change and a monic polynomial of degree $d$. \qed
\end{lemma}

\noindent In the case of formal groups over a perfect field of positive characteristic, this reduces to two familiar structural results:

\begin{corollary}
Every nonzero map of formal groups over a perfect field of positive charactersitic can be factored as an iterate of Frobenius and a coordinate change.\footnote{Incidentally, the Frobenius iterate appearing in the Weierstrass factorization of the multiplication-by-$p$ isogeny $p\co \G \to \G$ is another definition of the height of $\G$.} \qed
\end{corollary}

\begin{corollary}
A map of formal groups over a complete local ring with a perfect positive-characteristic residue field is an isogeny if and only if the kernel subscheme of the map is a divisor. \qed
\end{corollary}

This last result forms the headwaters of the connection we are seeking: isogenies are exactly the class of formal group homomorphisms whose kernels form subgroup divisors.  Again, we are seeking an assignment in the opposite direction, some special collection of endoisogenies naturally attached to prescribed kernel divisors.  As a first approximation to this goal, we drop the \emph{endo-} and aim to construct just \emph{isogenies} with this kernel property, a candidate for which is a theory of \textit{quotient groups}.

\begin{definition}
For $K \subseteq \G$ be a subgroup divisor, we define the quotient group $\G/K$ to be the formal scheme whose ring of functions is the equalizer
\begin{center}
\begin{tikzcd}
\sheaf O_{\G/K} \arrow{r} & \sheaf O_{\G} \arrow[shift left=0.4em, "\mu^*"]{r} \arrow[shift right=0.4em, "1 \otimes \eta^*"']{r} & \sheaf O_{\G} \otimes \sheaf O_K.
\end{tikzcd}
\end{center}
\end{definition}

\begin{lemma}[{\cite[Theorem 5.3.2-3]{StricklandFiniteSubgps}}]
The functor $\G/K$ is again a $1$--dimensional smooth commutative formal group. \qed
\end{lemma}

The inclusion of rings of functions determines an isogeny $q\co \G \to \G/K$ of degree $|K|$.  In this particular case, the induced pullback map $q^*$ of divisor schemes has an especially easy formulation:
\begin{lemma}
Pullback along the isogeny $q\co \G \to \G/K$ is computed by \[q^* \co D \mapsto D * K,\] where $*\co \Div_n^+ \G \times \Div_d^+ \G \to \Div_{nd}^+ \G$ is the convolution product of divisors on formal groups as described in \Cref{ProductMapOfDivisorSchemes}. \qed
\end{lemma}

In the case that $K$ is specified by a level structure, this admits a further refinement:
\begin{corollary}
Let $\ell\co A \to \G$ be a level structure parametrizing a subgroup divisor $K$.  The divisor pullback map can then be computed by the expansion \[q^* D = \sum_{a \in A} \tau_{\ell(a)}^* D,\] where $\tau_g\co \G \to \G$ is the translation by $g$ map. \qed
\end{corollary}

\begin{definition}\label{NormForFns}
This second construction can be upgraded to an assignment from \emph{functions} on $\G$ to \emph{functions} on $\G/K$, rather than just the ideals that they generate.  Specifically, we define the \textit{norm} of $\phi \in \sheaf O_{\G}$ along a level structure $\ell\co A \to \G$ by the formula \[N_\ell \phi = \prod_{a \in A} \tau_{\ell(a)}^* \phi.\]  An often-useful property of this norm construction is that if $\phi$ is a coordinate on $\G$, then $N_\ell \phi$ is a coordinate on $\G/K$~\cite[Theorem 5.3.1]{StricklandFiniteSubgps}.
\end{definition}

\begin{remark}
In general, the pullback map $q^*$ admits the following description: an isogeny $q\co C \to C'$ gives a presentation of $\sheaf O_C$ as a finite free $\sheaf O_{C'}$--module.  A function $\phi \in \sheaf O_C$ therefore begets a linear endomorphism $\phi \cdot (-) \in \operatorname{GL}_{\sheaf O_{C'}}(\sheaf O_C)$, and the determinant of this map gives an element of $q^* \phi \in \sheaf O_{C'}$.  Letting $\phi$ range, this gives a multiplicative (but not typically additive) map $q^*\co \sheaf O_C \to \sheaf O_{C'}$.  If a divisor $D$ is specified as the zero-locus of a function $\phi_D$, the divisor $q^* D$ is specified as the zero-locus of $q^* \phi_D$.
\end{remark}

Our last technical remark is that this definition of quotient does, indeed, have the suggested universal property:

\begin{lemma}[Third isomorphism theorem for formal groups, {\cite[Theorem 5.3.4]{StricklandFiniteSubgps}}]
If $f\co \G \to \widehat{\mathbb H}$ is any isogeny with kernel divisor $K$, then there is a uniquely specified commuting triangle
\begin{center}
\begin{tikzcd}
& \G \arrow["q"']{ld} \arrow["f"]{rd} \\
\G/K \arrow["g", "\simeq"']{rr} & & \widehat{\mathbb H}. \qed
\end{tikzcd}
\end{center}
\end{lemma}

We now use this Lemma, along with properties of the Lubin--Tate moduli problem, to associate endo-isogenies to subgroup divisors.  To begin, consider the multiplication--by--$p$ endo-isogeny of a Lubin--Tate group $\G$.  Since $\G$ is finite height, this map is an isogeny and the Lemma above gives rise to an isomorphism
\begin{center}
\begin{tikzcd}
& \G \arrow["q"']{ld} \arrow["p"]{rd} \\
\G/\G[p] \arrow["g", "\simeq"']{rr} & & \G.
\end{tikzcd}
\end{center}
In particular, this diagram shows that the quotient map $\G / \G[p]$ is \emph{again} a universal deformation, as witnessed by a preferred isomorphism to $\G$.  For a generic subgroup divisor $K \le \G$, we have access to the isogeny $q_K$ using the methods described above, but it is the magic of the Lubin--Tate moduli that furnishes us with replacements for $p$ and for $g$.  Notice first that the problem simplifies dramatically for the formal group at the special fiber: all subgroups of the special fiber formal group are of the form $p^j \cdot [0]$, and hence we can always use the Frobenius map to complete the desired triangle:
\begin{center}
\begin{tikzcd}
& \Gamma \arrow["q_{p^j[0]}"']{ld} \arrow["\Frob^j"]{rd} \\
\Gamma/(p^j\cdot[0]) \arrow["g_{p^j[0]}", "\simeq"']{rr} & & (\phi^j)^* \Gamma,
\end{tikzcd}
\end{center}
where $\phi\co k \to k$ is the Frobenius on coefficients and $\Frob\co \Gamma \to \phi^* \Gamma$ is the ``geometric Frobenius'', specified at the level of formal group \emph{laws} by the equation $\Frob(x) = x^p$ and \[(x +_\Gamma y)^p = x^p +_{\phi^* \Gamma} y^p.\]

\begin{lemma}[{\cite[Section 12.3]{AHSHinfty}}]
Let $G$ be an infinitesimal deformation of a finite height formal group $\Gamma$ to a complete local ring $R$.  Associated to a subgroup divisor $K \le G$, there is a commuting triangle
\begin{center}
\begin{tikzcd}
& G \arrow["q_K"']{ld} \arrow["P_K"]{rd} \\
G/K \arrow["g_K", "\simeq"']{rr} & & \psi_H^* \G
\end{tikzcd}
\end{center}
for a map $\psi_K\co \Spf R \to (\moduli{fg})^\wedge_\Gamma$ and $\G$ the Lubin--Tate universal deformation of $\Gamma$.
\end{lemma}
\begin{proof}
The main content of the deformation theory of finite height formal groups, recounted in \Cref{SectionMfgSmallScales}, is that there is a natural correspondence between the following two kinds of deformation data:
\[\left\{\begin{tikzcd}[column sep=0.7em]
\Gamma \arrow{d} & i^* \Gamma \arrow{l} \arrow{rd} \arrow[equal, "\alpha"]{r} & j^* G \arrow{r} \arrow{d} & G \arrow{d} \\
\Spec k & & \Spec R/\m \arrow["i"']{ll} \arrow["j"]{r} & \Spf R
\end{tikzcd}
\right\}
\leftrightarrow
\left\{\begin{tikzcd}[column sep=0.7em]
G \arrow{rd} \arrow[equal, "\beta"]{r} & \psi^* \G \arrow{d} \arrow{r} & \G \arrow{d} \\
& \Spf R \arrow["\psi"]{r} & (\mathcal M_{\mathbf{fg}})^\wedge_\Gamma
\end{tikzcd}
\right\}.\]
Accordingly, to construct the map $\psi_K$ from the Lemma statement, we need only exhibit $G/K$ as belonging to a natural diagram of the sort at left.  This is exactly what the Frobenius discussion above accomplishes:
\todo{Did you get the direction here right?}
\begin{center}
\begin{tikzcd}
\Gamma \arrow{d} & (\phi^j)^* \Gamma \arrow["\Frob^j"']{l} \arrow{rd} \arrow[equal, "g_{p^j[0]}"]{r} & (j^* G)/p^j\cdot[0] \arrow{r} \arrow{d} & G/K \arrow{d} \\
\Spec k & & \Spec k \arrow["\phi^j"']{ll} \arrow["j"]{r} & \Spf R.
\end{tikzcd}
\end{center}
Transferring this to a diagram of the sort at right, this gives the map $\psi_K$ and the isomorphism $g_K$, and the map $P_K$ is constructed as their composite.
\end{proof}

Applying this Lemma to the universal case gives the following result:

\begin{corollary}
There is a unique sequence of maps \[P^{\Sigma_{p^k}}\co \G \times \Sub_{p^k} \G \to \psi_{\Sigma_{p^k}}^* \G\] determined by the following properties:
\begin{enumerate}
    \item The kernel of $P^{\Sigma_{p^k}}$ is exactly the universal subgroup.
    \item (\textit{Deformation of Frobenius}:) At the special fiber, $P^{\Sigma_{p^k}}$ reduces to the $k${\th} Frobenius iterate.
    % (There is \emph{one} of these in each $\star$-isomorphism class.)
    % \item There is there a family of commuting triangles
    % \begin{center}
    % \begin{tikzcd}
    % & \G \arrow["N_K"']{ld} \arrow["q_K"]{rd} \\
    % \G / K \arrow["{g_K \quad \simeq}"]{rr} & & \psi_K^* \G
    % \end{tikzcd}
    % \end{center}
    % indexed on subgroups $K \subseteq \G$.
    \qed \todo{Prove me?}
\end{enumerate}
\end{corollary}

Amazingly, this is exactly the behavior of the power operation for $E$--theory when applied to $\CP^\infty$.  The $\Sigma_{p^k}$--power operation map in Morava $E$--theory \[E^0 \CP^\infty \otimes E^0 B\Sigma_{p^k} \from E^0 \CP^\infty\] becomes the following map of formal schemes:
\begin{center}
\begin{tikzcd}
\Spf (E^0 \CP^\infty \otimes E^0 B\Sigma_{p^k} / I_{\Tr}) \arrow{r} \arrow[equal]{d} & \Spf E^0 \CP^\infty \arrow[equal]{d} \\
\CP^\infty_E \times \Sub_{p^k} \CP^\infty_E \arrow["P^{\Sigma_{p^k}}"]{r} & \CP^\infty_E,
\end{tikzcd}
\end{center}
and this becomes an $E^0$--algebra map (i.e., a map over $(\moduli{fg})^\wedge_\Gamma$) when the bottom map is factored as \[\CP^\infty_E \times \Sub_{p^k} \CP^\infty_E \xrightarrow{P^{\Sigma_{p^k}}} \psi_{\Sigma_{p^k}}^* \CP^\infty_E \to \CP^\infty_E.\]  This map has exactly the prescribed kernel, and by its very nature as a \emph{power operation} it reduces to the Frobenius on the special fiber.  Finally, this identification is definitionally compatible with the map sending a power operation to its constituent sum of unstable operations, i.e., the map from the natural $E_\infty$ context to the unstable context.

\begin{remark}[{\cite[1.4.2.3]{CCO}}]
There is a useful result about isogenies that more topologists out to be aware of, although it doesn't tie in directly to any of our exposition here.  Nonetheless, we record its statement: let $R$ be a (nice) complete local ring, and let $\G$ and $\G'$ be two $p$--divisible groups over $R$.  There is an injection \[\Isog_R(\G, \G') \into \Isog_{R/\m}(\G, \G').\]
\end{remark}











\subsection*{$H_\infty$ orientations}

Having worked through enough of the underlying algebra, we now return to our intended topological application of studying $H_\infty$ orientations of Morava $E$--theory by $MUP$.  There are two reduction theorems, due to McClure, that lighten our workload from an infinite number of conditions to check to merely two conditions:

\begin{theorem}[{\cite[Proposition VIII.7.2]{BMMS}}]
Let $E$ and $F$ be $H_\infty$ ring spectra with $F$ $p$--local, and let $f\co E \to F$ be a map of ring spectra in the homotopy category.  Then $f$ is furthermore an $H_\infty$ ring map if and only if the following equation is satisfied: \[f \circ P^{\Sigma_p}_E = P^{\Sigma_p}_F \circ f^{\sm p}_{h\Sigma_p}. \qed\]
\end{theorem}

\begin{theorem}
\citeme{Find a Lawson citation for this last claim.}
Let $E$ be an $E_\infty$ ring spectrum and let $X$ be a space such that $E^{\Susp^\infty_+ X}$ is a wedge of copies of $E$.  The map \[\iota^* \otimes \Delta^*\co \widetilde E^* X^{\sm G}_{hG} \to \widetilde E^* X^{\sm G} \oplus \widetilde E^*(X \sm BG_+)\] is then injective.\footnote{The actual statement of McClure's result~\cite[Proposition VIII.7.3]{BMMS} has several additional hypotheses: $\pi_* E$ is taken to be even-concentrated and free over $\Z_{(p)}$, $X$ has homology free abelian in even dimensions and zero in odd dimensions, and $X$ and $E$ are both taken to be finite type.  Although this is the theorem cited in the source material~\cite[Section 4]{Ando}~\cite[Proof of Proposition 6.1]{AHSHinfty}, the version that has the weak hypotheses that we require only appeared in print much later.} \qed
\end{theorem}

\begin{corollary}[{\cite[pg.\ 271]{AHSHinfty}}]
\todo{You could prove this or mention where the character theory reduction happens.  This is also a really disgusting presentation of this rather simple diagram}
A ring map $x\co MUP \to E_\Gamma$ is $H_\infty$ if and only if the internal power operations commute: \[x \circ \mu^{\Sigma_p}_{MUP} \circ \Delta = \mu^{\Sigma_p}_{E_\Gamma} \circ x^{\sm p}_{h\Sigma_p} \circ \Delta \in E_\Gamma^0 T(\L \otimes \rho \downarrow \CP^\infty \times BC_p). \qed\]
\end{corollary}

We now expand the algebraic condition that this last Corollary encodes.  The power operation for $MUP$ was described in \Cref{AjAndBjAreInTheFGLSubring}, where we found that it applies the norm construction to $f$ for the universal $\Z/p$--level structure on $\G$.  The cyclic power operation on Morava $E$--theory was determined in the previous section to act by pullback along the deformation of Frobenius isogeny associated to the same universal level structure.  This condition is important enough to warrant a name:

\begin{definition}\label{NormCoherentDefn}
A coordinate $\phi\co \G \to \A^1$ on a Lubin--Tate group $\G$ is said to be \textit{norm-coherent} when for all subgroups $K \subseteq \G$, we have that $N_K \phi$ and $\psi_K^* \phi$ are related by the isomorphism $g_K$.
\end{definition}

\begin{corollary}
The subset of those orientations which are $H_\infty$ correspond exactly to those coordinates which are norm-coherent, and it suffices to check the norm-coherence condition just for the universal $\Z/p$--level structure. \qed
\end{corollary}

This characterization is already somewhat interesting, but it will only be truly interesting once we have found examples of such coordinates.  Our actual main result is that such coordinates are remarkably (and perhaps unintuitively) common.  In order to set up the statement and proof of this result, we consider the following somewhat more general situation:

\begin{definition}
A line bundle $\L$ on $\G$ together with a level structure $\ell\co A \to \G$ induces a line bundle $N_\ell \L$ on $N_\ell \G$ according to the formula \[N_\ell \L = \bigotimes_{a \in A} \tau_{\ell(a)}^* \L.\]  This line bundle interacts with the norm-coherence triangles well: $N_\ell \L = g_\ell^* \psi_\ell^* \L$.  However, the operations on individual sections can be different: a section $s \in \Gamma(\L \downarrow \G)$ is said to be \textit{norm-coherent} when for any choice of level structure we have $N_\ell s = g_\ell^* \psi_\ell^* s$.
\end{definition}

\begin{example}
In particular, functions $\phi\co \G \to \A^1$ can be thought of as sections of the trivial line bundle $\sheaf O_{\G}$, and coordinates can be thought of as trivializations of the same trivial line bundle.  We also remarked in \Cref{NormForFns} that the property of being a coordinate is stable under the operations on both sides of the norm-coherence equation.\footnote{This is reflected in topology as the assertion that the two power operations give two Thom classes for the regular representation, which must therefore differ by a unit.}
\end{example}

\begin{theorem}\label{AndosAlgebraicTheorem}
\citeme{Ando's thesis and also Peterson-Stapleton}
Let $\L$ be a line bundle on a Lubin--Tate formal group $\G$, and let $s_0$ be a section of $\L_0$, the restriction of $\L$ to the formal group at the special fiber.  Then there exists exactly one norm-coherent section $s$ of $\L$ itself which restricts to $s_0$.
\end{theorem}
\begin{proof}[Proof sketch]
The first observation is that the norm-coherence condition can be made sense of after reducing modulo any power of the maximal ideal in the Lubin--Tate ring, and that the resulting condition is trivially satisfied in the case where the entire maximal ideal is killed.  We then address the problem inductively: given a norm-coherent section modulo $\m^j$, we seek a norm-coherent section modulo $\m^{j+1}$ extending this.  By picking \emph{any} lift extending this, we can test the norm-coherence condition and produce an error-term that measures its failure to hold.  In the specific case of the canonical subgroup $\G[p] \le \G$, this error term is contained in the Lubin--Tate ring, and so we can modify our lift by subtracting off this error term.  This perturbation has no effect on the ``norm'' side of the norm-coherence condition, and it has a linear effect on the ``$\psi_{[p]}$'' side, so it cancels with the error term to give a section which is norm-coherent \emph{only when tested against the canonical subgroup}.

One can already show that the unicity clause applies: there is only one section $s$ of $\L$ which reduces to $s_0$ and which is norm-coherent for the canonical subgroup $\G[p]$.  In order to conclude that $s$ is truly norm-coherent, one shows that $N_\ell s$ satisfies its own $[p]$--norm-coherence condition (for the new canonical subgroup $(\G / \ell)[p] \le \G / \ell$), forcing it to agree with $\psi_\ell^* s$.
\end{proof}

\begin{corollary}\label{AHSHinftyResultForEthy}
Every orientation $s_0\co MUP \to K_\Gamma$ extends uniquely to a diagram
\begin{center}
\begin{tikzcd}
MUP \arrow["s_0"]{rd} \arrow["s"]{d} \\
E_\Gamma \arrow{r} & K_\Gamma,
\end{tikzcd}
\end{center}
where $s$ is an $H_\infty$ ring map. \qed
\end{corollary}

\begin{example}[{\cite[Section 2.7]{Ando}, cf.\ \Cref{ArtinHasseExponential}}]
The usual coordinate on $\G_m$ with $x +_{\G_m} y = x + y - xy$ satisfies the norm-coherence condition.  We compute directly in the case of the canonical subgroup:
\begin{align*}
(p = 2) & & N_{[2]}(x) & = x(x +_{\G_m} 2) \\
& & & = x (x + 2 - 2x) \\
& & & = 2x - x^2 = 2^*(x), \\
(p > 2) & & N_{[p]}(x) & = \prod_{j=0}^{p-1} (x +_{\G_m} (1 - \zeta^j)) \\
& & & = \prod_{j=0}^{p-1} (x + (1 - \zeta^j) - x(1 - \zeta^j)) \\
& & & = \prod_{j=0}^{p-1} (1 - \zeta^j(1 - x)) \\
& & & = (1 - (1 - x)^p) = p^*(x).
\end{align*}
This is exactly the computation that $g_{[p]}^* \psi_{[p]}^*(x)$ and $N_{[p]}(x)$ agree for this choice of $x \in \sheaf O_{\G_m}$.
\end{example}

The generality with which we approached the proof of \Cref{AndosAlgebraicTheorem} not only clarifies which operations apply to the objects under consideration\footnote{This is meant in contrast to Ando's original proofs~\cite[Section 2.6]{Ando}, where he deals only with coordinates and uses uncomfortable composition operations.}, but it also applies naturally to the other kinds of orientations discussed in \Cref{ChapterSigmaOrientation}.

\begin{theorem}
Orientations $MU[6, \infty) \to E_\Gamma$ which are $H_\infty$ correspond to norm-coherent cubical structures.  If $\height{\Gamma} \le 2$, orientations $M\String \to E_\Gamma$ which are $H_\infty$ correspond to norm-coherent $\Sigma$--structures. \qed
\end{theorem}

\begin{example}[{\cite[Sections 15-16]{AHSHinfty}}]
Let $C_0$ be an elliptic curve over a perfect field of positive characteristic.  The Serre--Tate theorem says that the infinitesimal deformation theory of $C_0$ is naturally isomorphic to the infinitesimal deformation theory of its $p$--divisible group $C_0[p^\infty]$, and we let $C$ be the universally deformed elliptic curve.  Our discussion of norm-coherence for Lubin--Tate groups can be repeated almost verbatim for elliptic curves, and we note further that level structures on $\widehat C$ inject into level structures on $C$.

We can apply these observations to $E_{\widehat C}$, the Morava $E$--theory for the formal group $\widehat C$, considered as an elliptic spectrum.  The natural orientation $MU[6, \infty) \to E_{\widehat C}$ from \Cref{EllipticSpectraAreOriented} is determined by the natural cubical structure on $\sheaf I(0)$, which by \Cref{Theta3IsTrivial} is \emph{uniquely} specified by the elliptic curve.  Our main new observation, then, is that the $N_\ell$ and $\psi_\ell^*$ constructions both convert this to a cubical structure on $C / \ell$, and hence are forced to agree by unicity.  In turn, it follows that the $\sigma$--orientation is a map of $H_\infty$ rings.
\end{example}



% \begin{remark}
% Power operations in Tate $K$--theory vs this story.
% \end{remark}


\todo{This MO conversation looks interesting: http://chat.stackexchange.com/transcript/message/29465746\#29465746.}
\todo{Bousfield's paper \textit{On $\lambda$--rings and the $K$--theory of infinite loopspaces} claims to have applications to Mahowald--Thompson type resolutions.}
\todo{Don Davis has a \textit{Handbook of Algebraic Topology} chapter on unstable $v_1$--periodic homotopy of spheres.}
\todo{Baker's POWER OPERATIONS AND COACTIONS IN HIGHLY COMMUTATIVE HOMOLOGY THEORIES seems like a nice place to learn about these things. He advertises some interaction with the traditional context story, which is appealing, and he mostly treats the case of ordinary homology, which we probably ought to spend a section on.}
\todo{FPFP Example 4.6 is a nice summary of a main result of HKR character theory.  Compare also Example 4.8.}
\todo{If you can figure out what Hood was doing with Dyer--Lashof operations in $\HFtwo$ and the logarithm/exponential coordinates, that'd be pretty cool to include.  (Neil has a Steenrod Algebra note that brushes against this.)}
\todo{Describe the action by $GL_n(\Z_p)$. (Hint at the action by $M_{n \times n}(\Z_p)$ with $\det \ne 0$.)}

















\section{The spectrum of modular forms}\label{ConstructionOfTMFSection}

\begin{center}
\textbf{\Large This is a paste of a talk from another venue. It needs reworking.}
\end{center}

This is a service talk.  One of the goals of this conference is to discuss Serre duality on derived stacks, and in order to explore this idea we require some interesting examples of derived stacks.  We have been continually interested in the moduli stack of elliptic curves, and this is the stack we will seek a topological enrichment for.  To begin: what is a topological enrichment?

\begin{definition}
Let $f\co \stack N \to \moduli{fg}$ be a flat, representable morphism of stacks.  A \textit{topological enrichment} of $\stack N$ is a sheaf of $E_\infty$ ring spectra $\sheaf O$ on $\stack N$ such that\todo{Talk about the complex-oriented cohomology of $\CP^1$.} \[\pi_n \circ \sheaf O \cong \begin{cases} f^* \omega^{\otimes k} & \text{if $n = 2k$ is even}, \\ 0 & \text{if $n$ is odd}. \end{cases}\]
\end{definition}

\begin{remark}
To gain intuition for the above, ignore the structure ring spectrum part and recall that a Landweber-flat even-periodic cohomology theory $R$ is determined by a flat map $\Spec R_0 \to \moduli{fg}$ from an affine, together with a preferred lift to $\moduli{fgl}$.  In this case, we have \[\pi_2 R = [\S^2, R] = [\CP^1, R] = \widetilde R^0 \CP^1 = \frac{\ker R^0 \CP^\infty \to R^0(*)}{(\ker R^0 \CP^\infty \to R^0(*))^{\otimes_{R^0 \CP^\infty} 2}},\] which is the cotangent space of $R^0 \CP^\infty$ or the invariant $1$--forms on $R^0 \CP^\infty$.\footnote{The identification of these two things is an indirect consequence of the discussion of logarithms from Luca's talk.}  This explains the condition about homotopy groups.
\end{remark}

\begin{remark}
To gain intuition about what a topological enrichment gains us, consider a flat map $\Spec R \to \stack N$.  We recover a complex-orientable homology theory $R_0(-)$ in the same way as Landweber's theorem, and its associated formal group is classified by the composite to $\moduli{fg}$.  Given an \'etale map $\Spec R \to \stack N$ and a topological enrichment $\sheaf O$, this homology theory has a chosen lift to an $E_\infty$ ring spectrum: $\sheaf O(R)$.
\end{remark}

Fix a fixed map $f\co \stack N \to \moduli{fg}$, this has its own associated moduli problem of topological enrichments of $f$.

\begin{theorem}[Goerss--Hopkins--Miller; Behrens; Lurie]
The moduli of topological enrichments of $\moduli{ell}$ is contractible.
\end{theorem}

\noindent The cohomology theories manufactured from this sheaf are examples of \textit{elliptic spectra}, which are triples $(E, C, \phi)$ of an even-periodic cohomology theory $E$, a elliptic curve $C$, and an isomorphism $\phi$ between $C^\wedge_0$ and the formal group associated to $E^0 \CP^\infty$.

We aren't going to prove this theorem---the details are too thick to completely present in an hour, even assuming much more topological technology---but I hope to give enough indications of the proof that you get the gist and are able to slot in the relevant topological inputs as you learn about them.  Our approach is to work locally on $\moduli{ell}$: divide it up over primes by passing to the $p$--completion, then divide the $p$--complete moduli itself into two further regions by using the following result:

\begin{lemma}
The $p$--divisible group of an elliptic curve is either formal of height $2$ or an extension of an \'etale $p$--divisible group of height $1$ by a formal group of height $1$. \qed
\end{lemma}
% \begin{proof}
% The Riemann--Manin symmetry condition states that the quasi-isogeneous components of a Dieudonn\'e module associated to an abelian variety are Cartier self-symmetric.  Then, an abelian variety of dimension $d$ has $p$--divisible group of height $2d$.
% \end{proof}

\noindent Elliptic curves over a $p$--complete ring with connected $p$--divisible groups (i.e., those of the first form) are called \textit{supersingular}, whereas those of the second form are called \textit{ordinary}.  The most important fact about supersingular elliptic curves is that they are reasonably uncommon:

\begin{lemma}
The supersingular locus of $\moduli{ell}$ is $0$--dimensional.  In fact, it is the zero-locus of a polynomial of degree $\lfloor (p-1)/12 \rfloor + \{0, 1, 2\}$. \qed
\end{lemma}

We write $i\co \moduli{ell}^{\ord} \to \moduli{ell}$ for the open inclusion of the ordinary locus.  We then plan to recover a topological enrichment by constructing the pieces of the following pullback:\footnote{This is an incarnation of the chromatic fracture square.  The top-right node is the $K(2)$--local component, the bottom-left is the $K(1)$--local component, and the bottom-right is the gluing data: the $K(1)$--localization of the $K(2)$--local component.}
\begin{center}
\begin{tikzcd}
& \sheaf O_{\top} \arrow{r} \arrow{d} & (\sheaf O_{\top})^\wedge_{\moduli{ell}^{\ss}} \arrow{d} & \sheaf O^{\ss} \arrow[equal]{l} \\
\sheaf O^{\ord} \arrow[equal]{r} & i_* i^* \sheaf O_{\top} \arrow{r} & i_* i^* \left((\sheaf O_{\top})^\wedge_{\moduli{ell}^{\ss}}\right).
\end{tikzcd}
\end{center}






\subsection*{The supersingular locus}

Our task in this section is to define $\sheaf O^{\ss}$ on $\widehat{\moduli{ell}^{\ss}}$, the supersingular part of the topological enrichment.  In order to make this definition, we need to specify its behavior on formal \'etale affines.  Since the moduli is itself $0$--dimensional, these are exactly the formal affine covers of the deformation spaces of the supersingular curves in the larger moduli $\moduli{ell}$.  The following arithmetic result gives us a crucial reduction:

\begin{theorem}[Serre--Tate]
The map $\moduli{ell} \to \moduli{pdiv}(2)$ is formally \'etale.\footnote{In general, the Serre--Tate theorem states that $\moduli{ab}^{d} \to \moduli{pdiv}(2d)$ is formally \'etale.} \qed
\end{theorem}
\begin{lemma}
The deformation theory of a connected $p$--divisible group of height $d$ as a $p$--divisible group is isomorphic to the deformation theory of the associated formal group of height $d$ as a formal group. \qed
\end{lemma}

This reduces us to a case from JD's talk: the inclusion of a deformation neighborhood of a height $2$ point on the moduli of formal groups, which he showed had an associated homology theory called \textit{Morava $E$--theory}.  A \emph{very} extravagant application of the tools from Ben's talk yields the following theorem, essentially owing to the very nice (i.e., formally smooth) deformation space and very nice (i.e., formally smooth) space of operations:

\begin{theorem}[Goerss--Hopkins--Miller]
Let $\Gamma$ be a finite height formal group over a perfect field.  The moduli of topological enrichments of $(\moduli{fg})^\wedge_\Gamma$ is homotopy equivalent to $B\Aut \Gamma$.  These operations are distinguished by their behavior on the cohomology of $\CP^\infty$. \qed
\end{theorem}

In particular, this gives a canonical topological enrichment of $\widehat{\moduli{ell}^{\ss}}$: it is the pullback of the Goerss--Hopkins--Miller sheaf along the Serre--Tate map \[\widehat{\moduli{ell}^{\ss}} = \coprod_{\text{supersingular $C$}} (\moduli{ell})^\wedge_C \xrightarrow{\text{f.\'e.}} \coprod_{\text{supersingular $C$}} (\moduli{pdiv}(2))^\wedge_{C[p^\infty]} \xleftarrow{\cong} \coprod_{\text{supersingular $C$}} (\moduli{fg})^\wedge_{\widehat C}.\]

\begin{remark}
This buys you more than just a bouquet of Morava $E$--theories, or even the global sections \[\sheaf O^{\ss}\left(\widehat{\moduli{ell}^{\ss}}\right) = \prod_{\text{supersingular $C$}} E_{\widehat C}^{h\Aut C}.\]  For instance, the moduli $\moduli{ell}^{\ss}(N)$ of supersingular elliptic curves $C$ equipped with specified isomorphisms $C[N] \cong (\Z/N)^{\times 2}$ forms an \'etale cover of $\moduli{ell}^{\ss}$ whenever $p \nmid N$, and hence this sheaf produces spectra $\TMF(N)^{\ss} = \sheaf O^{\ss}(\moduli{ell}^{\ss}(N))$ satisfying $(\TMF(N)^{\ss})^{hGL_2(\Z/N)} \simeq \TMF^{\ss}$.
\end{remark}







\subsection*{The ordinary locus}

Before addressing the ordinary locus in earnest, where our goal is to manufacture a \emph{lot} of cohomology theories, we spend a moment thinking about a particular example: $p$--adic $K$--theory, a familiar cohomology theory whose formal group looks like one associated to an ordinary elliptic curve.  Using the Goerss--Hopkins--Miller theorem stated above, $p$--adic $K$--theory appears as the $E_\infty$ ring\footnote{$K$--theory can be given the structure of an $E_\infty$ ring spectrum directly from the geometry of vector bundles, and it is a further theorem that these two structures agree.} of sections assigned to the \'etale cover \[\Spf \Z_p \to (\moduli{fg})^\wedge_{\G_m}.\]  Using $\Aut \G_m = \Z_p^\times$, the second half of the Goerss--Hopkins--Miller theorem endows this with a $\Z_p^\times$--indexed family of $E_\infty$ ring maps, denoted $\psi^k$ for $k \in \Z_p^\times$.  Their effect on the cohomology of $\CP^\infty$ is given by
\begin{align*}
\psi^k\co K_p^0 \CP^\infty & \to K_p^0 \CP^\infty, \\
x & \mapsto 1 - (1 - x)^k,
\end{align*}
where $x \in K_p^0 \CP^\infty$ is the ``usual'' power series generator.

On the other hand, the $E_\infty$ ring structure itself gives another map \[P^p\co K_p^0 \to K_p^0 B\Sigma_p,\] the $p${\th} power operation.  There is a calculation $K_p^0 B\Sigma_p = K_p^0\{1, \theta\}$, and in these terms the effect of the $p${\th} power operation on the cohomology of $\CP^\infty$ is
\begin{align*}
P^p\co K_p^0 \CP^\infty & \to K_p^0 \CP^\infty \otimes K_p^0\{1, \theta\}, \\
x & \mapsto (1 - (1 - x)^p) \otimes \theta.
\end{align*}
This extra operation describes an extension of the $\Z_p^\times$ action on $K_p$ to an action of the monoid $\Z_p$.\footnote{An alternative description of the effect of the power operation is that it encodes quotienting by an order $p$ subgroup of the formal group associated to $E_\Gamma$, and it so happens that $\G_m[p]$ is the \emph{unique} such subgroup for $\G_m$.}

In general, the $p$--adic $K$--theory of an $E_\infty$ ring spectrum also carries a $\Z_p$--action (and perhaps a little more, if the homotopy groups are not torsion-free), called a \textit{$\theta$--algebra}.  A different version of the Goerss--Hopkins--Miller theorem gives a reverse-engineering tool that converts information about $\theta$--algebras into information about the space of possible ordinary $E_\infty$ ring spectra yielding them on evaluation of $p$--adic $K$--theory, or the mapping space between two such.  A tool from JD's talk lets us guess which algebra should be associated to the $p$--adic $K$--theory of the global sections of $\sheaf O^{\ord}$, pictured in the first pullback square below:

\begin{center}
\begin{tikzcd}
\cdots \arrow{r} & \Spf W_1 \arrow{r} \arrow{d} & \Spf V^\wedge_\infty \arrow{r} \arrow{d} & \Spf \Z_p \arrow["{\text{f.\'e.}}"]{d} \\
\cdots \arrow{r} & \moduli{ell}^{\ord}(p^1) \arrow{r} & \moduli{ell}^{\ord} \arrow{r} & \moduli{fg}.
\end{tikzcd}
\end{center}

\noindent Moreover, $V^\wedge_\infty$ has a natural structure as a $\theta$--algebra: the interesting map $\psi^p$ acts by \[\psi^p\co (C, \eta\co \G_m \xrightarrow\cong \widehat C) \mapsto \left(
\begin{tikzcd}
\G_m[p] \arrow{r} \arrow{d} & \G_m \arrow["p"]{r} \arrow["\eta"]{d} & \G_m \arrow[densely dotted, red, "\eta^{(p)}"]{d} \\
C[p]\arrow{r} & C \arrow["p"]{r} & C^{(p)}
\end{tikzcd}
\right).\]  Unfortunately, this $\theta$--algebra is not nice enough to apply the Goerss--Hopkins--Miller theorem.  At this point, it becomes convenient to work at $p \ge 5$ for simplicity, where introducing a formal $(\Z/p)$--level structure fixes this:
\begin{lemma}[Igusa]
For $p \ge 5$, the moduli $\moduli{ell}^{\ord}(p)$ is \emph{affine}. \qed
\end{lemma}
\begin{corollary}
The associated $\theta$--algebra $W_1$ has vanishing Goerss--Hopkins--Miller obstruction groups, hence realizes uniquely to an ordinary $E_\infty$ ring spectrum $\TMF(p)^{\ord}$, and the action of $(\Z/p)^\times$ on the level structure enhances to a coherent $(\Z/p)^\times$--action on $\TMF(p)^{\ord}$. \qed
\end{corollary}

We define $\TMF^{\ord}$, our candidate for $\Gamma(\sheaf O^{\ord})$, to be the $(\Z/p)^\times$--fixed points of $\TMF(p)^{\ord}$, and indeed its $p$--adic $K$--theory is $V^\wedge_\infty$.  More than this, it turns out that the $\theta$--algebra associated to any formal \'etale affine open of $\moduli{ell}^{\ord}$ has a unique realization \emph{as an algebra under $\TMF(p)^{\ord}$}, and maps between such also lift uniquely---and this gives us the desired sheaf $\sheaf O^{\ord}$.  This is a common strategy: find a topological enrichment of an affine cover of your stack of interest, then descend it to the stack itself.





\subsection*{Gluing data}

The last thing we have to do to construct the pullback square is to manufacture a map of sheaves \[i_* i^* \sheaf O_{\top} \to i_* i^* \left((\sheaf O_{\top})^\wedge_{\moduli{ell}^{\ss}}\right).\]  This is rather similar to the construction of $\sheaf O^{\ord}$ itself: we construct a candidate map $\TMF^{\ord} \to (\TMF^{\ss})^{\ord}$ of global sections, and then we use this to control the map of sheaves using relative Goerss--Hopkins obstruction theory.

The main result that marries algebra to topology are the following two results about $(\TMF^{\ss})^{\ord}$.  The first is that $(\TMF^{\ss})^{\ord}$ counts as an elliptic spectrum:

\begin{lemma}
There is an elliptic curve $C^{\alg}$ over an affine $\Spf ((V^\wedge_\infty)^{\ss})$ such that $(\TMF^{\ss})^{\ord}$ is an elliptic spectrum for this curve. \qed
\end{lemma}

\noindent This gives two candidates for a $\theta$--algebra structure on the $p$--adic $K$--theory of $\TMF^{\ss}$: there is the $\theta$--algebra structure coming from the geometric map $\Spf ((V^\wedge_\infty)^{\ss}) \to \Spf V^\wedge_\infty$, and there is the $\theta$--algebra structure coming from the shear fact that $(\TMF^{\ss})^{\ord}$ is an ordinary $E_\infty$ ring spectrum.

\begin{theorem}
The natural $\theta$--algebra structure on $\Spf ((V^\wedge_\infty)^{\ss})$ induced by the map $\Spf ((V^\wedge_\infty)^{\ss}) \to \Spf V^\wedge_\infty$ agrees with the Goerss--Hopkins--Miller $\theta$--algebra structure on $\pi_* (\TMF^{\ss})^{\ord}$. \qed
\end{theorem}

\noindent This is to be read as a recognition theorem for the $\theta$--algebra structure on the topological object $(\TMF^{\ss})^{\ord}$: it matches the algebraic model.  Once this is established, the Goerss--Hopkins--Miller obstructions can be shown to vanish, and it follows that the above map lifts to a map of the $E_\infty$ rings of global sections, and then one proceeds to produce the map of sheaves by further applications of obstruction theory.

Arithmetic fracture is dealt with similarly, but it is \emph{far} simpler.  Because $\Q \otimes \TMF$ has a smooth $\Q$--algebra as its homotopy, the obstructions governing the version of Goerss--Hopkins--Miller for commutative $H\Q$--algebras vanish, letting us lift algebraic results into homotopy theory wholesale.






\subsection*{Variations on these results}

\begin{remark}
At the prime $3$, the proof of Igusa's theorem needs amplification, but the statement remains the same and the rest of the story goes through smoothly.
\end{remark}

\begin{remark}
At the prime $2$, two further things go wrong: one must pass to the Igusa cover $\moduli{ell}^{\ord}(4)$ before it becomes affine, but then the Galois group of this cover is $C_2$, which has infinite cohomological dimension at $2$.  Appealing to the equivalence $KO = KU^{hC_2}$, one works with $2$--adic \emph{real} $K$--theory instead, which somehow pre-computes the Galois action.
\end{remark}

\begin{remark}
There is another way to construct $\TMF^{\ord}$ at low primes, given by a complex consisting of two $E_\infty$ cells attached to $\S$.  The way this is done, essentially, is by constructing a complex whose $p$--adic $K$--theory matches the expected value: first it must have the right dimension, and then the action of $\theta$ must be corrected.
\end{remark}

\begin{remark}
There is an analogous (and much easier) picture for the moduli of forms of the multiplicative group: any ordered pair of puncture points in $\mathbb A^1$ can be used to give $\mathbb P^1$ the unique structure of a group with identity at $\infty$, and the associated formal group is classified by a map $\moduli{\mathbb G_m} \to \moduli{fg}$; there is an equivalence $\moduli{\mathbb G_m} \simeq BC_2$; and $KU$ forms the global sections of a topological enhancement of $\Spec \Z \to \moduli{fg}$ which descends using the complex-conjugation action to $BC_2 \to \moduli{fg}$.
\end{remark}

\begin{remark}
With some effort\footnote{Actually, a significant chunk of the trouble is already present in the details of this construction, since dealing with the gluing data requires working with an algebraized curve, which is only classified by a map to the compactified moduli.}, the construction of $\sheaf O_{\top}$ outlined here extends to the compactified moduli $\overline{\moduli{ell}}$ where Weierstrass curves with nodal singularities are allowed, i.e., where $\Delta$ is \emph{not} inverted (as in $y^2 + xy = x^3$).  The resulting global sections yields a spectrum $\Tmf$, which is \emph{not} a periodic ring spectrum.  The connective truncation of that spectrum is denoted $\tmf$, and it is expected to arise as the global sections of a topological enrichment of a stack of generalized cubics, i.e., where cuspidal singularities are also allowed (as in $y^2 = x^3$).  The existence proof for a topological enrichment of $\moduli{cub}$ remains elusive.
\end{remark}




\subsection*{Descent on homotopy}

One of the main upsides of finding one of these topological enrichments is that it comes a equipped with a spectral sequence computing the homotopy of its global sections, coming from recovering $\sheaf O(\stack N)$ as the homotopy limit of finer and finer covers of $\stack N$:
\begin{lemma}
For $\sheaf O$ a topological enrichment of an appropriate map $\stack N \to \moduli{fg}$, there is a spectral sequence \[E_2^{s, t} = H^s(\stack{N}; \pi_t \circ \sheaf O) \Rightarrow \pi_{t-s} \Gamma(\stack{N}; \sheaf O). \qed\]
\end{lemma}

\begin{lemma}
This spectral sequence is isomorphic to the $MU$--Adams spectral sequence for $\sheaf O(\stack N)$.
\end{lemma}
\begin{proof}[Main observation]
Consider the \v{C}ech complex associated to the affine cover $\moduli{Weier} \to \moduli{ell}$.  We claim that the complex making up the $E_1$--term of the descent spectral sequence is isomorphic to the complex making up the $E_1$--term of the $MU$--Adams spectral sequence.  To illustrate, we compute the first two terms of each and compare them.
\begin{enumerate}
    \item Consider the pullback diagram of stacks
    \begin{center}
    \begin{tikzcd}
    \Spec A \arrow{r} \arrow{d} & \moduli{fgl} \arrow{d} \\
    \moduli{ell} \arrow{r} & \moduli{fg}.
    \end{tikzcd}
    \end{center}
    In the same stroke, this is also the pullback diagram computing $\Spec MU_* \TMF$.
    \item Now consider the pair of cubes of iterated pullbacks pictured in \Cref{HypercubesFigure}.  These compute the pullback of the cube in two different ways, producing an isomorphism $\moduli{quad.trans.} \cong (\moduli{fg} \times \moduli{ps}^{\mathrm{gpd}}) \times_{\moduli{fgl}} \moduli{Weier}$.
    \item[$n$.] The general case is similar, but requires stomaching iterated pullbacks in $n$--cubes. \qedhere
\end{enumerate}
\end{proof}

\begin{figure}
    \begin{center}
    \begin{tikzcd}[column sep=0em]
    & \begin{array}{c}(\moduli{fgl} \times \moduli{ps}^{\mathrm{gpd}}) \\ \times_{\moduli{fgl}} \\ \moduli{Weier} \end{array} \arrow{rr} \arrow{ld} & & \moduli{Weier} \arrow{dd} \arrow{ld} \\
    \moduli{fgl} \times \moduli{ps}^{\mathrm{gpd}} \arrow{rr} \arrow{dd} & & \moduli{fgl} \arrow{dd} \\
    & & & \moduli{ell} \arrow{ld} \\
    \moduli{fgl} \arrow{rr} & & \moduli{fg},
    \end{tikzcd}
    \end{center}
    \begin{center}
    \begin{tikzcd}[column sep=0em]
    & \moduli{quad.trans.} \arrow{rr} \arrow{dd} & & \moduli{Weier} \arrow{dd} \arrow{ld} \\
    & & \moduli{fgl} \\
    & \moduli{Weier} \arrow{rr} \arrow{ld} & & \moduli{ell} \arrow{ld} \\
    \moduli{fgl} \arrow{rr} & & \moduli{fg}. \arrow[leftarrow, crossing over]{uu}
    \end{tikzcd}
    \end{center}
    \caption{Two expressions of the same cubical pullback.}\label{HypercubesFigure}
\end{figure}

\noindent We now appeal to Katharine's and Dominic's talks in order to compute $\pi_* TMF[1/6]$.  Since $2$ and $3$ are both inverted, we can use scaling and translation to complete both the cube and the square to replace an arbitrary Weierstrass curve with \emph{unique} one of the form $y^2 = x^3 + c_4 x + c_6$, with $\Delta = -24(4c^3_4 + 27c^2_3)$ and in fact the map \[\Spec \Z[c_4, c_6, \Delta^{-1}][1/6] \to \moduli{ell} \times \Spec \Z[1/6]\] is an equivalence of \todo{graded?} stacks.\todo{$c_4^3 - c_6^2 = 1728\Delta$}  Since the quasicoherent sheaf cohomology of affines is always amplitude $0$, this spectral sequence is concentrated on the $0$--line, and we recover \[\pi_* \TMF[1/6] \cong \Z[c_4, c_6, \Delta^{-1}][1/6].\]

\begin{remark}
This is all a rather elaborate way of recovering the homotopy of the \emph{complex-orientable} ring spectrum $\TMF[1/6]$.
\end{remark}

\todo{Cite Johan and Andr\'e}
\todo{The $\tmf$ sheaf exists: $\moduli{ell}$ has a $j$ map to $\mathbb A^1$, $\overline{\moduli{ell}}$ has a $j$ map to $\mathbb P^1 = \mathbb A^2 \setminus \{0\} / \Gm$, and that can be completed (by enlarging the moduli) to a $j$ map to $\mathbb A^2 / \Gm$.  \'Etale maps miss this extra bad point, so this is mostly just fussing with gluing data, and then only in the rational situation.}



\subsection*{---------------------------------------------------------------------------}



\subsection*{Recollections}

This talk is the first mention of $\TMF$, the spectrum of \textit{topological modular forms}.  This is precisely where it made its historical appearance: it lies between the classical theory of chromatic homotopy theory described so far and the wide-open world of derived algebraic geometry about to be discussed.  Here are some classical ingredients that enter into $\TMF$:

\begin{lemma}
The ring of modular forms is $\Z[E_4, E_6]$.  In fact, the map $\Spec \Z[E_4, E_6][1/6] \to \moduli{ell}$ is an isomorphism away from $6$.
\end{lemma}
\begin{proof}[Mnemonic]
The inertia groups of $\moduli{ell}$ are of order $2$ and $3$, so inverting these primes causes $\moduli{ell}$ to become a scheme rather than a stack.  Weierstrass presentation finishes the isomorphism claim.
\end{proof}



\begin{corollary}
There is a complex-oriented cohomology theory, $\TMF[1/6]$, determined by \[\TMF[1/6]_0(X) = MUP_0(X) \otimes_{MUP_0} \Z[E_4, E_6][1/6].\]
\end{corollary}
\begin{proof}
In particular, the isomorphism above is flat.  Apply Landweber flatness.
\end{proof}

This ring spectrum has a kind of universal property:
\begin{definition}
An elliptic spectrum $E = (E, C, \phi)$ is an even-periodic ring spectrum $E$, a choice of an elliptic curve $C$, and a choice of an isomorphism $\phi\co C^\wedge_0 \cong \CP^\infty_E$.  Such a spectrum is called \textit{flat} when the map $E_* \to \moduli{ell}$ classifying $C$ is flat.
\end{definition}

\begin{corollary}
If $E$ is a flat elliptic spectrum with $1/6 \in \pi_0 E$, then there is a natural map $\TMF[1/6] \to E$. \qed
\end{corollary}

\begin{remark}
Similar spectra were classically investigated by different people: Morava studied a form of $\TMF$ ``near the cusp'' in \textit{Forms of $K$--theory}, and Landweber--Ravenel--Stong defined a version of elliptic cohomology specifically for the Jacobi curve.
\end{remark}

Our goal now is to understand $\TMF$ as an object in derived algebraic geometry.  This could mean many different things, the most basic of which is that we would like to instantiate $\TMF$ as an $E_\infty$ ring spectrum, so that we can make use of the tools from Ben's talk.  This will be our main immediate concern, but we will be guided by a considerably more complicated apparatus: as modular forms (of a particular weight) appear as sections of a certain sheaf on $\moduli{ell}$, we will also eventually arrive at a definition of $\TMF$ itself as the global sections of a sheaf of $E_\infty$ ring spectra, defined on the \'etale site of $\moduli{ell}$.  This is quite trippy to think about, so for now we will just use it as inspiration.

Recall the following Lemma:




\noindent Consequentially, $(\moduli{ell})^\wedge_p$ decomposes into two components: a closed substack of those curves of the first type, called \textit{supersingular}, and an open complement of those curves of the second type, called \textit{ordinary}.  Our approach to constructing $\TMF$ will be by chromatic fracture: we will define $L_0 \TMF$, $\widehat L_1 \TMF$, and $\widehat L_2 \TMF$, as well as certain gluing data that assembles into the chromatic fracture cube to recover $\TMF$ itself.

In order to avoid \emph{significant} complications introduced by the geometry of $\moduli{ell}$ itself, we will continue to work with $6$ inverted, thereby deleting all of the inertia groups.






\subsection*{The chromatic components of $\TMF[1/6]$}

\subsection*{The $K(1)$--local component}

We will investigate the $K(1)$--local part of $\TMF$ first.  The idea is to use continuous $p$--adic $K$--theory as a replacement for $K(1)$, since $p$--adic $K$--theory is an $E_\infty$ ring, and so this way we can make use of tools from Ben's talk.

\begin{theorem}[McClure]
The homotopy of a $K(1)$--local commutative $K_p$--algebra spectrum $R$, such as $\widehat L_1 (K_p \sm E)$, carries an extra family of ring operations $\psi^k$ indexed on $k \in \Z_p$, as well as a ring map $\theta$,\footnote{If $E_*$ is torsion-free, then the last condition means that $\theta$ is redundant.} such that
\begin{align*}
\psi^1(x) & = x, &
\psi^k (\psi^{k'} x) & = \psi^{kk'}(x), &
\psi^k(x) \cdot \psi^{k'}(x) & = \psi^{k + k'}(x), &
\psi^p(x) & = x^p + p\theta(x). \qed
\end{align*}
\end{theorem}

Graded rings with this structure are called \textit{graded $\theta$--$(K_p)_*$--algebras}, or just $\theta$--algebras.  The homotopy of $E$ itself can be recovered from that of $R = \widehat L_1 (K_p \sm E)$ by taking fixed points for the $\Z_p^\times$--action, i.e., by an Adams spectral sequence.  We say that a $K(1)$--local $E_\infty$ ring spectrum $E$ \textit{models} a $\theta$--algebra $A_*$ when $K^\vee_p E = A_*$.

\begin{theorem}[{Goerss--Hopkins, $K(1)$--locally}]
Given a map of $\theta$--algebras $f\co A_* \to B_*$, the following Andr\'e--Quillen cohomology groups (internal to $\theta$--algebras) measure various obstructions:
\begin{center}
\begin{tabular}{c|cc}
& existence & uniqueness \\
\hline
a model $E$ for $A$ & $H^{s \ge 3}_\theta(A_*, \Loops^{s-2} A_*)$ & $H^{s \ge 2}_\theta(A_*, \Loops^{s-1} A_*)$ \\
a map $E \to F$ of models & $H^{s \ge 2}_\theta(A_*, \Loops^{s-1} B_*)$ & $H^{s \ge 1}_\theta(A_*, \Loops^s B_*)$.
\end{tabular}
\end{center}
Finally, given such a map $f$, there is a spectral sequence computing the homotopy groups of the $E_\infty$ mapping space: \[E_2^{s, t} = H^s_\theta(A_*, \Loops^{-t} B_*) \Rightarrow \pi_{-s-t}(E_\infty(E, F), f). \qed\]
\end{theorem}

These enhanced Andr\'e--Quillen cohomology groups can be computed using a Grothendieck-type spectral sequence, intertwining classical Andr\'e--Quillen cohomology groups for commutative rings with the extra task of checking compatibility with the $\theta$--algebra structure.  We won't need to do this so manually, and the following utility result will capture what we need:

\begin{lemma}\label{ObstructionLemma}
Suppose that $A_*$ is a torsion-free even-periodic $\theta$--algebra, that $B_*$ is a torsion-free $\theta$--$A_*$--algebra, that $A_*^{\Z_p^\times} / p$ is a formally smooth $\F_p$--algebra, that $A_* / p$ is ind-\'etale over $A_*^{\Z_p^\times}/p$, and lastly that $H^{s \ge 1}_c(\Z_p^\times; B_*) = 0$.  Then the following vanishing result holds: \[H^{s \ge 1}(A_*, \Loops^t B_*) = 0, \quad H^0(A_*, \Loops^{2t} B_*) = 0. \qed\]
\end{lemma}

Now we make an educated guess about what $K_p^\vee \TMF$ should be, based on \textit{From Spectra to Stacks}\todo{Get JD to talk about this}:
\begin{center}
\begin{tikzcd}[column sep=0.3em, row sep=0.3em]
\text{``$\Spf K_p^\vee \TMF(p^k)$''} \arrow{rr} \arrow{rd} \arrow{dd} & & \moduli{ell}^{\ord}(p^k) \arrow{dd} \arrow{rd} \\
& \text{``$\Spf K_p^\vee \TMF$''} \arrow[crossing over]{rr} & & \moduli{ell} \arrow{dd} \\
\Spf \Z_p \arrow{rr} \arrow{rd} & & \moduli{fg}^{\mathrm{mult}}(p^k) \arrow{rd} \\
& \Spf K_* \arrow{rr}[description]{\G_m} \arrow[crossing over, leftarrow]{uu} & & \moduli{fg}^{\mathrm{mult}},
\end{tikzcd}
\end{center}
where the front and right faces are meant to be pullbacks.  The front pullback is the moduli of elliptic curves equipped with a \emph{marking}, which is an isomorphism $\eta\co \G_m \xrightarrow{\cong} \widehat C$.  This turns out to be affine with ring of functions $V^\wedge_\infty$, and the moduli-theoretic interpretation of this ring endows it with a natural structure as a $\theta$--algebra:
\begin{itemize}
  \item For $k \in \Z_p^\times$, we send a marked elliptic curve to $\psi^k(C, \eta) = (C, [k] \circ \eta)$.
  \item For $p \in \Z_p$, we use the fact that the map $p\co \G_m \to \G_m$ is an endomorphism of $\G_m$ with kernel exactly $\G_m[p]$.  This tells us that the $p${\th} multiple of a marking factors through the $p${\th} multiplication map on the elliptic curve, to give a marking $\eta^{(p)}$ of $C^{(p)}$.  This defines the relevant map by $\psi^p(C, \eta) = (C^{(p)}, \eta^{(p)})$.
\end{itemize}

\todo[inline]{This description of what's going on in the back face is \emph{not} right.}
This turns out to be insufficient to construct $\widehat L_1 \TMF$, because this $\theta$--algebra does not satisfy the hypotheses of \Cref{ObstructionLemma}.  However, the back face will do: it turns out that $\moduli{ell}^{\ord}(p^k)$ is affine, and $\moduli{ell}^{\ord}(p)$ is an \'etale torsor over $\moduli{ell}^{\ord}$ with structure group $G = \Z_p^\times / 1 + p\Z_p$.\footnote{The proof is that $\moduli{ell}^{\ord}(p^k)$ embeds into the moduli of ordinary elliptic curves equipped with a $(p^k-1)$--jet, and there is a unique change of coordinates that transforms an $n$--jet to one with vanishing first three components in the standard Weierstrass coordinate.}  You can write down a (completely analogous) $\theta$--algebra structure on this ring, and this time the obstruction theory finds a foothold: there exists a modeling spectrum $\tmf(p)^{\ord}$, the Hurewicz map from $\theta$--algebras to $\End_{E_\infty} \tmf(p)^{\ord}$ is an isomorphism (as spaces), and so the Galois action lifts to an action on $\tmf(p)^{\ord}$.  We set $\widehat L_1 \TMF = (\tmf(p)^{\ord})^{hG}$, and because $|G|$ is coprime to $p$ we have $K_p^\vee \widehat L_1 \TMF = V^\wedge_\infty$.

\begin{remark}
Before moving on, we list what modifications are necessary to get this to work at the remaining primes:
\begin{itemize}
    \item At $3$, the main obstacle is in the analysis of the Igusa tower: showing that $\moduli{ell}^{\ord}(p) \to \moduli{ell}^{\ord}$ has the relevant properties is complicated considerably by not being able to complete the cube.
    \item At $2$, the above modifications are necessary, but it is $\moduli{ell}^{\ord}(4)$ that is affine.  However, $(\Z/4)^\times \cong C_2$ is \emph{not} of order coprime to $2$, and indeed has infinite cohomological dimension.  As is often the case in this kind of situation, the correction is to re-do all of the preceding theory with $KO = KU^{hC_2}$ in place of $KU$, which has this action already built-in.
\end{itemize}
\end{remark}





\subsection*{The $K(2)$--local component}

The definition of the $K(2)$--local component of $\TMF$ actually proceeds along much the same lines.  The main fact we need to know about $\moduli{ell}$ is:






\noindent It is also useful to know the following fact about $K(n)$--localization:
\begin{lemma}
If $M$ is a $BP$--module spectrum, then $\pi_* \widehat L_n M = M[v_n^{-1}]^\wedge_{I_n}$. \qed
\end{lemma}

We might thus expect $\widehat L_2 \TMF$ to be assembled out of the formal neighborhood of the supersingular locus.  The Serre--Tate theorem says that these formal neighborhoods are isomorphic to the formal neighborhoods of the associated $p$--divisible groups in their moduli, and the formal neighborhoods of \emph{connected} $p$--divisible groups (which is the supersingularity condition) are isomorphic to the formal neighborhoods in $\moduli{fg}$.  We are therefore interested in the Morava $E$--theories associated to supersingular elliptic curves.  An amplification of the Goerss--Hopkins obstruction theorem described above also shows that Morava $E$--theories exist as $E_\infty$ ring spectra (owing, most essentially, to their formally smooth homotopy rings), and the only $E_\infty$ maps among them are controlled by the action of the Morava stabilizer.  Coupling these together, we are done.




\subsection*{Chromatic fracture}

We now seek to understand the diagram
\begin{center}
\begin{tikzcd}
\TMF[1/6]^\wedge_p \arrow{r} \arrow{d} & \widehat L_2 \TMF[1/6] \arrow{d} \\
\widehat L_1 \TMF[1/6] \arrow{r} & \widehat L_1 \widehat L_2 \TMF[1/6].
\end{tikzcd}
\end{center}
The right-hand vertical arrow is just the localizing unit associated to $\widehat L_1$, but the bottom horizontal arrow is more interesting: if we already had a spectrum $\TMF[1/6]^\wedge_p$, it would be $\widehat L_1$ applied to the localizing unit associated to $\widehat L_2$, but the whole point is that we do not.  Instead, we will define $\TMF[1/6]^\wedge_p$ by defining this bottom arrow.  The Goerss--Hopkins theorem gave us access to the space of $E_\infty$ maps off of $\TMF(p)^{\ord}$, but not $\TMF^{\ord}$, so our goal is to build a $G$--equivariant $E_\infty$ ring spectrum over $\widehat L_2 \TMF$, understand its $K(1)$--localization, produce a $G$--equivariant map of the desired form, and finally pass to fixed points.

\begin{definition}
The Galois cover $\moduli{ell}^{\ord}(p)$ came from considering markings, and the analogous structure that enhances to $\moduli{ell}^{\ss}$ is called a \textit{full level--$N$ structure}, which is an isomorphism $\eta\co (C_N)^2 \xrightarrow\cong C[N]$.\footnote{This marks the $p$--divisible group associated to $C$ rather than the formal group.}  The moduli of such elliptic curves with full level--$N$ structure forms a \emph{scheme} over $\Z[1/N]$ which is a $GL_2(\Z/N)$--Galois cover of $\moduli{ell} \times \Spec \Z[1/N]$, and we define the supersingular moduli with level by the pullback
\begin{center}
\begin{tikzcd}
\moduli{ell}^{\ss}(N) \arrow{r} \arrow{d} & \moduli{ell}(N)_p \arrow{d} & & & \moduli{ell}^{\ss}(N)^{\ord} \arrow{d} \arrow{r} & \moduli{ell}^{\ord} \arrow{d} \\
\moduli{ell}^{\ss} \arrow{r} & (\moduli{ell})_p, & & \moduli{ell}^{\ss}(N) \arrow{r} & \moduli{ell}^{\ss}(N)^{\alg} \arrow{r} & \overline{\moduli{ell}},
\end{tikzcd}
\end{center}
which we then algebraize and extend to the compactification, then define the ordinary part of the supersingular locus by intersection.  Using $\moduli{ell}^{\ss}(N)^{\ord}$, we then define a series of auxiliary pullbacks:
\begin{center}
\begin{tikzcd}[column sep=1em]
& \moduli{ell}^{\ss}(p)^{\ord} \arrow{rd} \\
\moduli{ell}^{\ss}(N,p)^{\ord} \arrow{rr} \arrow[red, "/GL_2(\Z/N)"]{ru} \arrow{d} & & \moduli{ell}^{\ord}(p) \arrow{d} & \moduli{ell}^{\ord}(N, p)^{\ns} \arrow[red]{r} \arrow{d} & \moduli{ell}^{\ord}(p)^{\ns} \arrow{d} & \moduli{ell}^{\ss}(N,p)^{\ord} \arrow{r} \arrow{d} & \moduli{ell}^{\ss}(p)^{\ord} \arrow{d} \\
\moduli{ell}^{\ss}(N)^{\ord} \arrow{rr} & & \moduli{ell}^{\ord} & \moduli{ell}(N)^\wedge_p \arrow{r} & (\moduli{ell})^\wedge_p & \moduli{ell}^{\ord}(N, p)^{\ns} \arrow{r} & \moduli{ell}^{\ord}(p)^{\ns}.
\end{tikzcd}
\end{center}
\end{definition}

\begin{lemma}[Remark 8.2]
The following are iterated pullback diagrams:
\begin{center}
\begin{tikzcd}
\Spf \widetilde W^{\ss} \arrow{r} \arrow{d} & \Spf W^{\ss} \arrow{r} \arrow{d} & \Spf W \arrow{r} \arrow{d} & \moduli{ell}^{\ord}(p^\infty) \arrow{d} \\
\moduli{ell}^{\ss}(N, p)^{\ord} \arrow{r} & \moduli{ell}^{\ss}(p)^{\ord} \arrow{r} & \moduli{ell}^{\ord}(p) \arrow{r} & \moduli{ell}^{\ord},
\end{tikzcd}
\end{center}
\begin{center}
\begin{tikzcd}[column sep=7em]
\Spf \widetilde W^{\ss} \arrow{r}[description]{/GL_2(\Z/N)} \arrow{dd} & \Spf W^{\ss} \arrow{r}[description]{/G} \arrow{dd} & \Spf ((V^\wedge_\infty)^{\ss}) \arrow{d} \arrow{r} & \moduli{ell}^{\ord}(p^\infty) \arrow{d} \\
& & \moduli{ell}^{\ss}(p)^{\ord} \arrow{r} \arrow{d} & \moduli{ell}^{\ord}(p) \\
\moduli{ell}^{\ord}(N, p)^{\ns} \arrow{r} & \moduli{ell}^{\ord}(p)^{\ns} \arrow{r} & (\moduli{ell}^{\ord})^{\ns}.
\end{tikzcd}
\end{center}
\end{lemma}

We make the definitions
\begin{align*}
\widehat L_2 \TMF(N) & := \sheaf O^{\top}_{K(2)}(\moduli{ell}^{\ss}(N)), &
\widehat L_1 \widehat L_2 \TMF(N, p) & = \widehat L_1 \widehat L_2 \TMF(N) \sm E_1^{h(1 + p\Z_p)}, &
\widehat L_1 \widehat L_2 \TMF(p) & = \widehat L_1 \widehat L_2 \TMF \sm E_1^{h(1 + p\Z_p)}.
\end{align*}
We make the computations
\begin{align*}
\pi_0 \widehat L_2 \TMF(N) & \cong \sheaf O_{\moduli{ell}^{\ss}(N)}, &
\pi_0 \widehat L_1 \widehat L_2 \TMF(N) & \cong O_{\moduli{ell}^{\ss}(N)}[u_1^{-1}]^\wedge_p,
\end{align*}
and we remark on the pullback square
\begin{center}
\begin{tikzcd}
\Spf \pi_0 (E \sm E_1^{h(1 + p\Z_p)}) \arrow{r} \arrow{d} & \moduli{fg}^{\mathrm{mult}}(p) \arrow{d} \\
\Spf \pi_0 E \arrow{r} & \moduli{fg}^{\mathrm{mult}}.
\end{tikzcd}
\end{center}
Using the first iterated pullback diagram from the Lemma, it follows that $\widehat L_1 \widehat L_2 \TMF(N, p)$ models $\widetilde W^{\ss}$.  Using the second iterated pullback square, we see that the other spectra are models for the other $\theta$--algebras pictured.  The Goerss--Hopkins theorem shows that there exists a unique $G$--equivariant map to the model $\widehat L_1 \TMF(p)$ for $W$.  Taking fixed points gives the desired chromatic attaching map.












The gluing data (remarks about $2$ and $3$):
  just computing the homotopy of $\widehat L_1 \widehat L_2 \TMF$ to get a guess at the moduli;
  look at 6.8--10 for $L_K(1) E_{\widehat C_{ss}}$;
  needs $\sheaf O^{\top}_{K(2)}$ to be a sheaf and needs to understand $\widehat L_1 \TMF \to \widehat L_1 \widehat L_2 \TMF$ as a map of $E_\infty$ $K(1)$--local elliptic spectra




\subsection*{Arithmetic fracture}

This is the easiest piece of the whole bunch.  Because $\pi_* L_{\Q} \tmf = \Q[a_4, a_6]$ is smooth, there are no Andr\'e--Quillen obstructions to constructing maps of commutative $H\Q$--algebras with signature \[L_{\Q} \tmf \to \mathbb A_{\mathrm{fin}} \otimes L_{\Q} \tmf.\]









\subsection*{The sheaf $\sheaf O^{\top}$ away from $6$}

$\TMF[1/6]$ as:
  the sections of a sheaf,
    as well as basic properties of homotopy sheaves,

enriching $\widehat L_1 \TMF$ to a sheaf:
  constructing $\Gamma(\Spec R; \moduli{ell}^{\ord})$ as $\widehat L_1 \TMF[1/6]$--algebras (pg.\ 163):
    if $E$ is $K(1)$--local elliptic classified by a flat map to $\moduli{ell}$, then $K^\wedge E = W$, where $\Spf W$ is the pullback of $\Spec E_0$ to $\moduli{ell}^{\ord}(p^\infty)$,
    $W$ has a unique $\theta$--algebra structure on it, and we say an $E_\infty$ structure on $E$ is $\theta$--compatible when these match,
    7.7 constructs a $(\Z/p)^\times$--equivariant spectrum $\tmf^{\ord}(p)_{K(1)}$,
    then checks that this has the right formal group,
  ;
enriching the gluing data to a map of sheaves:
  (pg.\ 172)
  ;
the universal property of the sheaf $\sheaf O^{\top}_p$:
  Proposition 8.5 (pg.\ 175)
  ;



\begin{remark}[Hopkins--Miller]
Remarks on $A_\infty$ and a different dissection of the moduli
\end{remark}

Extension to $\Tmf$ and to $\tmf$:
  singular and additive degenerations;
  the $\theta$--algebra structure in the $K(1)$--local analysis is complicated, seems solvable with work for multiplicative degeneration but not additive;
  the Witten genus and behavior at the cusp;
  



\subsection*{The descent spectral sequence}

Cohomology of the Hopf algebroid:
  emphasize that this is easier than studying the fracture square

$\pi_* \tmf[1/6] = \Z[1/6, a_4, a_6]$.

Remark on the horizontal vanishing line in the descent spectral sequence, Tate vanishing.

Remark on Stojanoska duality:
  maybe just make the statement for $\Tmf(2)$



JOHAN WANTS:
    $\TMF$ WITH LEVEL STRUCTURES;
    A MENTION OF THE PREDICTED MODULI UNDERLYING $\tmf$;
    A MENTION OF THE DESCENT SPECTRAL SEQUENCE FOR HOMOTOPY


















\section{Orientations by \texorpdfstring{$E_\infty$}{Eoo} maps}\label{JuvitopTalkSection}

A more modern take on the story of the $\sigma$--orientation passes directly through the algebra of $E_\infty$--ring spectra.  Though technically intensive, our reward for grappling with this will be the modularity of the $\String$--orientation, enriching \Cref{WittensTheoremForBU6} to the real setting.  Luckily, most of the basic ideas are classically familiar, centering on a particular functor \[\gl_1\co E_\infty\CatOf{RingSpectra} \to \CatOf{Spectra}.\]  This functor derives its name from two compatible sources: for one, its underlying infinite loopspace is the construction $GL_1$ described in \Cref{LectureThomSpectra}; and secondly, it participates in an adjunction
\begin{center}
\begin{tikzcd}[column sep=4em]
\CatOf{ConnectiveSpectra} \arrow[shift left=0.3\baselineskip, "\Susp^\infty_+ \Omega^\infty"]{r} & E_\infty\CatOf{RingSpectra} \arrow[shift left=0.3\baselineskip, "\gl_1"]{l}
\end{tikzcd}
\end{center}
analogous to the adjunction between the group of units and the group-ring constructions in classical algebra.  Its relevance to us is its participation in the theory of highly structured Thom spectra.  Let $j\co g \to \gl_1 \S$ be a map of connective spectra, begetting a map $J\co G \to \GL_1 \S$ of infinite loopspaces, where we have written $G = \Loops^\infty g$.
\begin{lemma}
\citeme{May? or ABGHR?}
The Thom spectrum of the map $BJ$ is presented by the pushout of $E_\infty$ rings\footnote{This is a kind of ``twisted group-ring'' construction.}
\begin{center}
\begin{tikzcd}
\Susp^\infty_+ \GL_1 \S \arrow["\Loops^\infty \Susp j"]{r} \arrow{d} & \Susp^\infty_+ \Loops^\infty \gl_1 \S / g \arrow{d} \\
\S \arrow{r} & MG. \qed
\end{tikzcd}
\end{center}
\end{lemma}

\begin{corollary}
\citeme{May, but also ABGHR}
There is a natural equivalence between the space of null-homotopies of the composite \[g \xrightarrow j \gl_1 \S \xrightarrow{\gl_1 \eta_R} \gl_1 R\] and the space of $E_\infty$ ring maps $MG \to R$, where $MG$ is the Thom spectrum of the stable spherical bundle classified by $J$.
\end{corollary}
\begin{proof}
Applying the mapping space functor $E_\infty(-, R)$ to the pushout diagram in the Lemma, we have a pullback diagram of mapping spaces:
\begin{center}
\begin{tikzcd}
E_\infty(\Susp^\infty_+ \GL_1 \S, R) & E_\infty(\Susp^\infty_+ \Loops^\infty \gl_1 \S / g, R) \arrow{l} \\
E_\infty(\S, R) \arrow{u} & E_\infty(MG, R) \arrow{l} \arrow{u}.
\end{tikzcd}
\end{center}
We can reidentify each of the three terms to get
\begin{center}
\begin{tikzcd}
\CatOf{Spectra}(\gl_1 \S, \gl_1 R) & \CatOf{Spectra}(\gl_1 \S / g, \gl_1 R) \arrow{l} \\
\{\gl_1 \eta_R\} \arrow{u} & E_\infty(MG, R) \arrow{l} \arrow{u},
\end{tikzcd}
\end{center}
hence $E_\infty(MG, R)$ appears as the fiber at $\gl_1 \eta_R$ of the restriction map, which coincides with the space of nullhomotopies as claimed.
\end{proof}

\begin{corollary}
\citeme{AHR}
The mapping set $E_\infty(Mj, R)$ is nonempty if and only if $\gl_1 \eta_R \circ j$ is null-homotopic.  If this is the case, then $E_\infty(Mj, R)$ is a torsor for $[\Susp g, \gl_1 R]$.
\end{corollary}

Ando, Hopkins, and Rezk have used this presentation to understand the mapping space $E_\infty(M\String, \tmf)$.  In this Appendix, we will use this same technology to understand the mapping space $E_\infty(M\Spin, KO_{(p)})$, which proceeds along entirely similar lines but is a \emph{considerably} simpler computation.\footnote{Ando, Hopkins, and Rezk also do $E_\infty(M\Spin, KO)$ as a warm-up computation~\cite[Section 7]{AHR}, and we are further $p$--localizing that result so as not to have to think about arithmetic fracture.  Working arithmetically globally should be an easy exercise for the reader.}  The approach to this computation is to mix the presentation above with chromatic fracture applied to the target:\todo{Put a pullback corner here.}
\begin{center}
\begin{tikzcd}
M\Spin \arrow{r} \arrow[bend left=15]{rr} \arrow{rrd} \arrow{rd} & KO_{(p)} \arrow{r} \arrow[crossing over]{d} & KO_p \arrow{d} \\
& \Q \otimes KO \arrow{r} & \Q \otimes KO_p.
\end{tikzcd}
\end{center}
So, we seek a pair of $E_\infty$ ring maps into the rationalization and the $p$--completion of $KO$ which agree on the $p$--local ad\`eles, which involves understanding not just the mapping spaces but also the pushforward maps between them.


\subsubsection{Rational orientations}

We begin with the two rational nodes in the pullback diagram.  As a first approximation to our goal, consider the problem of giving a complex orientation $MU \to \Q \otimes R$ of a rational ring spectrum $\Q \otimes R$.  There is an automatic such orientation granted by
\begin{center}
\begin{tikzcd}
MU \arrow[densely dotted, "D"]{r} \arrow{rd} & \Q \otimes R \\
\S \arrow{u} \arrow[crossing over]{ru} \arrow{r} & H\Q \arrow{u}
\end{tikzcd}
\end{center}
constructed out of the unit map $\S \to MU$, the unit map $\S \to \Q \otimes R$, the rationalization map $S \to \Q \otimes \S \cong H\Q$, and the standard additive orientation $MU \to H\Q$ of an Eilenberg--Mac Lane spectrum.  When $E_\infty(MU, T)$ is nonempty, it is a torsor for $[bu, \gl_1 T]$, and since we have a preferred orientation $D$ we thus have isomorphisms \[\pi_0 E_\infty(MU, \Q \otimes R) \xleftarrow{\cong} [bu, \gl_1 \Q \otimes R] \xleftarrow{\cong} [bu, \Q \otimes \gl_1 R] \xrightarrow{\cong} [\Q \otimes bu, \Q \otimes \gl_1 R],\] the last of which is specified by a sequence of rational numbers $(t_{2k})_{k \ge 1}$.  The role played by the sequence $(t_{2k})$ is to perturb the Thom class.

\begin{lemma}
Write $x$ for the Thom class of $\L$ on $\CP^\infty$ in $(\Q \otimes R)$--cohomology as furnished by the automatic orientation $D$.  The Thom class associated to some other orientation of $\Q \otimes R$ is tracked by a difference series $x / \exp_F(x)$, and the sequence $(t_k)$ above is expressed by $x / \exp_F(x) = \exp(\sum_k t_k/k! \cdot x^k)$.\todo{This is confused.}
\end{lemma}
\begin{proof}[Proof sketch]
Let $v^k\co S^{2k} \to BU$ be the $k${\th} power of the class $\L$, so that it comes from a restriction \[S^{2k} \to (\CP^\infty)^{\sm k} \xrightarrow{\L^{\boxtimes k}} BU.\]  The Thom class for this bundle comes from the top Chern class, which is the top coefficient in the product of total Chern classes applied to the individual bundles.  Following the usual formulas shows the map $v^k$ to behave on homotopy by multiplication by $(-1)^k t_k$.
\end{proof}

Now we move away from $MU$.  There are three directions for generalization: connective orientations, real orientations, and non-complex targets.
\begin{enumerate}
\item Rationally, the analysis of Ando--Hopkins--Strickland identifies $[BU\<2k\>, \Q \otimes R]$ with $k$--variate symmetric multiplicative $2$--cocycles over $R$, every one of which arises as $\delta^1$ repeatedly applied to a univariate series.  In homotopy theoretic terms, this means that every $MU\<2k\>$--orientation of a rational spectrum factors through an $MU$--orientation.
\item The cofiber sequence $kO \to kU \to \Susp^2 kO$ splits rationally, using the idempotents $\frac{1 \pm \chi}{2}$ on $kU$.  Accordingly, $MU$--orientations of rational spectra that factor through $MSO$--orientations have an invariance property under $\chi$: $-[-1](x) = x$, corresponding to the idempotent factor $+$.  This pattern continues for the characteristic series of connective orientations.
\item This same cofiber sequence and idempotent splitting also tells us that rational $KU$--cohomology classes in the image of $KO$--cohomology are $\chi$--invariant, i.e., they belong to the $-$ factor.
\end{enumerate}

Our main example is the usual orientation $MU \to KU$ that selects the formal group law $x + y - xy$.  This is associated to the difference Thom class $x / (e^x - 1) = x / \exp_{\G_m}(x)$.  To make this difference $[-1]$--invariant (and hence give a complex-orientation of $KO$), we use the averaged exponential class $(e^{x/2} - 1) - (e^{-x/2} - 1)$.\footnote{Incidentally, this is equal to $2\operatorname{sinh}(x/2)$.}  In turn, we use the Lemma to calculate the behavior on homotopy of the associated orientation:\footnote{This comes out of applying $d\log$ to the fraction.} \[\frac{x}{e^{x/2} - e^{-x/2}} = \exp\left(-\sum_{k=2}^\infty \frac{B_k}{k} \cdot \frac{x^k}{k!}\right).\]  Finally, we calculate the effect of the orientation on the second half of the factorization \[MSU \to M\Spin \to KO,\] again using the relevant idempotent, which has the effect of halving the coefficients in the characteristic series: $-\frac{B_k}{2k}$.\footnote{While we're here, you might want to observe that elements in $[bu, \gl_1 R]$ push forward to elements in $[bu, \gl_1 \Q \otimes R]$ which do not disturb the denominators of the elements $t_k$.  (On the other hand, the ``Miller invariant'' associated to a rational ring spectrum is \emph{zero}, because arbitrary elements in $[bu, \gl_1 \Q \otimes R]$ can completely destroy the denominators.)}

This discussion accounts for both $E_\infty(M\Spin, \Q \otimes KO)$ and $E_\infty(M\Spin, \Q \otimes KO_p)$: the set of rational characteristic series includes into the set of ad\`elic characteristic series as the subset with rational coefficients.




\subsubsection{Finite place orientations}\label{FinitePlaceOrientationsSubsection}
\newcommand{\spin}{\mathit{spin}}

We want now to understand $E_\infty(M\Spin, KO_p)$ and its map to $E_\infty(M\Spin, \Q \otimes KO_p)$.  Here's the initial set-up:
\begin{center}
\begin{tikzcd}
\spin \arrow{r}[description]{j} & \gl_1 \S \arrow{r} \arrow{rd}[description]{\gl_1 \eta_{KO_p}} & Cj \arrow[densely dotted, "A"']{d} \\
& & \gl_1 KO_p.
\end{tikzcd}
\end{center}
We are looking to understand the space of filler diagrams $A$ (i.e., vertical maps with choice of homotopy of the precomposite to $\gl_1 \eta_{KO_p}$).  Notice first that there is a natural cofiber sequence to be placed on the bottom row:
\todo{I don't like the placing of this $A$. I want it to indicate a choice of filler.}
\begin{center}
\begin{tikzcd}[column sep=1em]
\spin \arrow{r}[description]{j} \arrow[red]{rd} & \gl_1 \S \arrow{r} \arrow[red]{d} \arrow{rd}[description]{\gl_1 \eta_{KO_p}} & Cj \arrow[densely dotted, "A"']{d} \\
& \Susp^{-1} \Q/\Z \otimes \gl_1 KO_p \arrow{r} & \gl_1 KO_p \arrow{r} & \Q \otimes \gl_1 KO_p \arrow{r} & \Q/\Z \otimes \gl_1 KO_p.
\end{tikzcd}
\end{center}
There is a canonical red vertical lift of $\gl_1 \eta_{KO_p}$ since $\gl_1 \S$ is a torsion spectrum, and this precomposes with $j$ to give another vertical map.  Notice now that selecting a filler triangle $A$ gives a commuting square with choice of homotopy and that $[\gl_1 \S, \Q \otimes \gl_1 KO_p] = 0$, and hence we would get a natural map (and natural homotopy) off of the homotopy cofibers:
\begin{center}
\begin{tikzcd}[column sep=1em]
\spin \arrow{r}[description]{j} \arrow{rd} & \gl_1 \S \arrow{r} \arrow{d} \arrow{rd}[description]{\gl_1 \eta_{KO_p}} & Cj \arrow[densely dotted, "A"' near start]{d}[description]{B} \arrow{r} & b\spin \arrow{r} \arrow{rd} \arrow[densely dotted]{d}[description]{C}& b\gl_1 \S \arrow{d} \\
& \Susp^{-1} \Q/\Z \otimes \gl_1 KO_p \arrow{r} & \gl_1 KO_p \arrow{r} & \Q \otimes \gl_1 KO_p \arrow{r} & \Q/\Z \otimes \gl_1 KO_p,
\end{tikzcd}
\end{center}
where $C$ is a map making the triangle it belongs to commute.  This all gives a function assigning $A$ to $B$ and $A$ to $C$ (and, in fact, the latter assignment factors through the former).

In order to show nonconstructively that the set of $A$s is nonempty, we might try to discern that $\gl_1 \eta_{KO_p} \circ j \in [\spin, \gl_1 KO_p]$ is zero by demonstrating something about the mapping set $[\spin, \gl_1 KO_p]$ itself.  We proceed by a sequence of quite improbable steps, beginning with the following Theorem original to Ando--Hopkins--Rezk:
\begin{theorem}[{\cite[Theorem 4.11]{AHR}}]
Let $R$ be a $E(d)$--local $E_\infty$ ring spectrum, and set $F$ to be the fiber \[F \to \gl_1 R \to L_d \gl_1 R.\]  Then $\pi_* F$ is torsion and $F$ satisfies the coconnectivity condition $F \simeq F(-\infty, d]$. \qed
\end{theorem}

\noindent It follows that $\gl_1 KO_p \to L_1 \gl_1 KO_p$ is a $1$--connected map, and hence \[[\spin, \gl_1 KO_p] = [\spin, L_1 \gl_1 KO_p].\]  In fact, we can even pass to the $K(1)$--localization, if we digress for a moment to introduce Rezk's logarithmic cohomology operation.

\begin{lemma}[{\cite[Theorem 1.1]{Kuhn}}]
For each $d \ge 1$ there is a functor $\Phi_d\co \CatOf{Spaces}_{*/} \to \CatOf{Spectra}$ which commutes with finite limits, is insensitive to upward truncation, and which evaluates on infinite loopspaces to give $\Phi_d(\Loops^\infty X) = \widehat L_d X$.\footnote{Importantly, $\Phi_d$ does \emph{not} care about the actual infinite loopspace structure on $\Loops^\infty X$, just that it has \emph{some} lift to a spectrum $X$.}\footnote{There is also a version of this theorem for $d = 0$, but since rational localization has no periodic behavior the results as not nearly as striking.} \qed
\end{lemma}

\begin{definition}[{\cite[Section 3]{RezkLogarithm}}]
The natural equivalence $(\GL_1 R)[1, \infty) \to (\Loops^\infty R)[1, \infty)$ gives rise to a map $\ell$ as in the diagram
\begin{center}
\begin{tikzcd}
& \Phi_d (\GL_1 R)[1, \infty) \arrow["\simeq"]{r} & \Phi_d (\Loops^\infty R)[1, \infty) \\
\gl_1 R \arrow{r} \arrow[bend left=15, "\ell_d" near end]{rr} & \widehat L_d \gl_1 R \arrow["\simeq"]{r} \arrow[equal, crossing over]{u} & \widehat L_d R \arrow[equal, crossing over]{u} .
\end{tikzcd}
\end{center}
\end{definition}

\begin{remark}
Applying the logarithm to the corners in the height $1$ chromatic fracture square yields the following identification:
\begin{center}
\begin{tikzcd}
& L_1 \gl_1 R \arrow{rr} \arrow{dd} & & \widehat L_1 R \arrow{dd} \\
L_1 \gl_1 R \arrow[crossing over]{rr} \arrow{dd} \arrow[equal]{ru} & & \widehat L_1 \gl_1 R \arrow{ru}[description]{\ell_1} \\
& \widehat L_0 R \arrow{rr} & & \widehat L_0 \widehat L_1 R \\
\widehat L_0 \gl_1 R \arrow{rr} \arrow{ru}[description]{\ell_0} & & \widehat L_0 \widehat L_1 \gl_1 R \arrow{ru}[description]{\widehat L_0 \ell_1} \arrow[crossing over, leftarrow]{uu} .
\end{tikzcd}
\end{center}
The front and back faces are connected by logarithms of \emph{different} heights---or, equivalently, the bottom horizontal arrow of the back face is \emph{twisted} from the usual chromatic fracture presentation of $L_1 R$.  The identification of this map is the usual sticking point in this approach.
\end{remark}

\begin{theorem}[{\cite[Theorem 1.9]{RezkLogarithm}}]
For $R$ a $K(1)$--local $E_\infty$ ring with $\pi_0 R$ torsion--free, the map $\pi_0 \ell_1\co \pi_0 R^\times \to \pi_0 R$ is given by the formula\footnote{The analogue of this formula for $E_\Gamma$ (but not an arbitrary $K(d)$--local $E_\infty$ ring spectrum) is given in \cite[Subsection 1.10]{RezkLogarithm}.} \[\ell_1(x) = \frac{1}{p} \log\left(\frac{x^p}{\psi^p x}\right) = \sum_{k=1}^\infty \frac{p^{k-1}}{k} \left(\frac{\theta(x)}{x^p}\right)^k. \qed\]
\end{theorem}

\begin{corollary}
The natural map $L_1 \gl_1 KO_p \to \widehat L_1 \gl_1 KO_p$ is a connective equivalence.
\end{corollary}
\begin{proof}
We specialize the above square to $R = KO_p$:
\begin{center}
\begin{tikzcd}
& & & KO_p \arrow{dd} \\
L_1 \gl_1 KO_p \arrow{rr} \arrow{dd} & & \widehat L_1 \gl_1 KO_p \arrow{ru}[description]{\ell_1} \\
& L_0 KO_p[4, \infty) \arrow{rr} & & L_0 KO_p \\
L_0 \gl_1 KO_p \arrow{rr} \arrow{ru}[description]{\ell_0} & & L_0 \widehat L_1 \gl_1 KO_p. \arrow{ru}[description]{\ell_1} \arrow[crossing over, leftarrow]{uu}
\end{tikzcd}
\end{center}
The behavior of the back horizontal map is determined by Rezk's formula for the logarithm.  \textbf{It acts by some nonzero number in every positive degree,}\todo{Justify this.} hence the fiber has the form $\prod_{k=-\infty}^0 \Susp^{4k-1} H\Q$.  Since the front face is a fiber square, this is also a calculation of the fiber of the map in the Lemma statement.\footnote{As a corollary of this same method, the Rezk logarithm for $R = KU^\wedge_p$ gives an equivalence $\gl_1 KU^\wedge_p[3, \infty) \to KU^\wedge_p[3, \infty)$.  This was previously known by nonconstructive methods to Adams and Priddy~\cite[Corollary 1.4]{AdamsPriddy}.}
\end{proof}

As a consequence, we have identifications \[\gl_1 \eta_{KO_p} \circ j \in [\spin, \gl_1 KO_p] \cong [\spin, L_1 \gl_1 KO_p] \cong [\spin, \widehat L_1 \gl_1 KO_p].\]  A direct application of the Rezk logarithm replaces $\widehat L_1 \gl_1 KO_p$ with $KO_p$, and the $K(1)$--localization of $\spin$ recovers $\Susp^{-1} KO_p$.  Altogether, this identifies $\gl_1 \eta_{KO_p} \circ j$ with a point in the mapping set $[\Susp^{-1} KO_p, KO_p]$---and we mark this as a point where we would like to understand the space of $KO$--operations.

We claim also that the kernel of the assignment $A \mapsto C$ is easy to understand: two fillers $A$ are related by an element of $[b\spin, \gl_1 KO_p]$, and their corresponding $C$s are related by the corresponding element of $[b\spin, \Q \otimes \gl_1 KO_p]$.  This set is rational, hence factors through the rationalization of $[b\spin, \gl_1 KO_p]$ where it must already be null, and hence it is a torsion element of $[b\spin, \gl_1 KO_p]$.  Meanwhile, the same argument as above identifies \[[b\spin, \gl_1 KO_p] = [KO_p, KO_p],\] which we again mark as a point where we would like to understand the space of $KO$--operations.  In particular, if we were to find the group of degree-preserving $KO$--operations to be torsion-free, then the assignment $A \mapsto C$ would be \emph{injective}.

We would like to understand the behavior of $C$ on homotopy based on some data about $A$.  This serves two purposes: there is the necessary condition that the triangle formed by $C$ and the canonical map $b\spin \to \Q / \Z \otimes \gl_1 KO_p$ commute, and then also the composite \[b\spin \xrightarrow{C} \Q \otimes \gl_1 KO_p \to (\gl_1 (\Q \otimes KO_p))[1, \infty)\] describes the map into the ad\`elic component.  In order to gain access to $C$, first notice that we can postcompose $B$ with the localization map off of $\gl_1 KO_p$ as in \Cref{MainAHRDiagram}.\footnote{Importantly, and differently from what every source says, this isn't a map of cofiber sequences and so the back second vertical map does not have to exist.}  This gives a new map $B'\co KO_p \to KO_p$---another reason to understand $KO$--operations.

We are now in a position to compute the action of $C$ on a homotopy class in $\pi_* b\spin$ by chasing through the following steps:
\begin{enumerate}
    \item We push such a class forward to $\widehat L_1 b\spin \simeq KO_p$ along the localization map.
    \item We then pull it back to $\widehat L_1 Cj \simeq KO_p$ along $KO_p \xrightarrow{1 - \psi^c} KO_p$, which acts by multiplication by $(1 - c^k)$ on $\pi_{4k}$.
    \item We push it down along $B'$ to $\widehat L_1 \gl_1 KO_p \simeq KO_p$, which acts by an unknown factor.
    \item We include it into the rational component of $\Q \otimes \widehat L_1 gl_1 KO_p$, using the fact that $\pi_* \widehat L_1 \gl_1 KO_p$ is torsion--free.
    \item Finally, we pull it back to $\Q \otimes \gl_1 KO_p$ along the logarithm $\ell_1$, which acts by multiplication by $(1 - p^{k-1})$ using Rezk's $K(1)$--local formula.\footnote{The formula for the logarithm in nonzero degrees comes from thinking of the logarithm as a \emph{natural transformation} and applying it to the mapping set $\ell\co \gl_1 KO^0(S^{2n}) \to KO^0(S^{2n})$.}
\end{enumerate}
The effect of this sequence of steps is \[t_{4k} = (1 - c^k)^{-1} b_{4k} (1 - p^{k-1})^{-1},\] where $t_{4k}$ and $b_{4k}$ are the effects on $\pi_{4k}$ of the maps $C$ and $B'$ respectively.  In the course of this proof, we are using the fact that division in the ring $\Z_p$ is unique when it is possible---the more responsible-looking equation to write is \[b_{4k} = (1 - c^k) t_{4k} (1 - p^{k-1}).\]

\begin{sidewaysfigure}
\centering
\begin{tikzcd}[column sep=0em]
& & & \widehat L_1 \S \arrow{rr} \arrow[equal]{d} & & KO_p \arrow[equal]{d} \arrow["1 - \psi^c"]{rr} & & KO_p \arrow[equal]{d} \\
& & & \widehat L_1 \gl_1 \S \arrow{rr} & & \widehat L_1 Cj \arrow[densely dotted, "B'" near start]{dd} \arrow{rr} & & \widehat L_1 b\spin \\
\spin \arrow{rr}[description]{j} \arrow{rrdd} & & \gl_1 \S \arrow{rr} \arrow{dd} \arrow{ru} \arrow{rrdd}[description]{\gl_1 \eta_{KO_p}} & & Cj \arrow{ru} \arrow[densely dotted, "A"' near start]{dd}[description]{B} \arrow[crossing over]{rr} & & b\spin \arrow{ru} \arrow[crossing over]{rr} \arrow[bend left=20]{rrdd} & & b\gl_1 \S \arrow{dd} \\
& & & & & \widehat L_1 \gl_1 KO_p \arrow{rr} & & \Q \otimes \widehat L_1 \gl_1 KO_p \\
& & \Susp^{-1} \Q/\Z \otimes \gl_1 KO_p \arrow{rr} & & \gl_1 KO_p \arrow{ru} \arrow{rr} & & \Q \otimes \gl_1 KO_p \arrow{ru} \arrow{rr} \arrow[densely dotted, leftarrow, crossing over]{uu}[description]{C} & & \Q/\Z \otimes \gl_1 KO_p.
\end{tikzcd}
\caption{A diagram showing the interconnections among the main components of the $p$--primary part of the Ando--Hopkins--Rezk argument.}\label{MainAHRDiagram}
\end{sidewaysfigure}

Now, finally, the diagonal map $b\spin \to \Q/\Z \otimes \gl_1 KO_p$ becomes relevant.  To check the commutativity of the triangle with $C$, we need only compare the results of the composite on homotopy since the map $C$ targets a rational spectrum and hence is determined its effect on homotopy.  The following invariance property makes this map accessible:\footnote{It is also possible to compute the effect of this map on homotopy using the $S^1$--transfer.  This is the subject of a paper by Miller~\cite{MillerBernoulliNos}, after which the Miller invariant is named, and also the subject of further research by Baker and company~\cite{BCGHRW}.}

\begin{theorem}[{\cite[Proposition 3.15 and Corollary 3.16]{AHR}}]
For any $A_\infty$ orientation $\phi\co MU \to R$ of an $A_\infty$ ring spectrum $R$, the denominators of the characteristic series associated to $\Q \otimes \phi$ compute the behavior of the map $\pi_* BU \to \Q / \Z \otimes GL_1 R$. \qed
\end{theorem}

\begin{corollary}
The numbers $t_{4k}$ describing the effect of $C$ satisfy the congruences \[t_{4k} \equiv -\frac{B_k}{2k} \pmod{\Z}.\]
\end{corollary}
\begin{proof}[Proof sketch]
The Todd orientation $MU \to KU$ is known to be $A_\infty$~\cite[Theorem V.4.1]{EKMM}, and the characteristic series of the Todd orientation has coefficients $B_k$.  The extra division by $2$ is picked up by studying the map $\pi_* BSU \to \pi_* B\Spin$ and the map $\pi_* KO \to \pi_* KU$.
\end{proof}

We have thus identified the legal fillers $C$ as those sequences of rational numbers $t_{4k}$ satisfying conditions:
\begin{enumerate}
    \item $t_{4k}$ has the correct denominators: for $k \ge 1$, $t_{4k} \equiv -B_k/(2k) \pmod{\Z}$.
    \item $b_{4k}$ is the effect on homotopy of some map $B'\co KO_p \to KO_p$.
\end{enumerate}


\subsubsection{Stable $KO$ operations}
\newcommand{\cts}{\mathrm{cts}}

We have identified three points where we want to understand the collection of stable $KO$ operations.  Although much of the main text of this book has been concerned with this sort of subject, this does not appear to be so immediately accessible: we want operations rather than cooperations, and $KO$ is \emph{not} a complex-orientable ring spectrum.  It is close to one, though, and we gain access to it through familiar approximation.

The easy initial calculation is $K^\vee K = \cts(\Z_p^\times, \Z_p)$, the ring of $\Z_p$--valued functions\footnote{Not homomorphisms!} on $\Z_p^\times$ which are continuous for the adic topologies on the domain and the target.  This comes out of the stable cooperations of Landweber flat homology theories discussed in \Cref{DefnChromaticHomologyThys}, where we showed that $E_\Gamma$ has cooperations given by the ring of functions on the pro-\'etale group scheme $\Aut \Gamma$.  For $\Gamma = \G_m$, this group scheme $\Aut \G_m$ is constant at $\Z_p^\times$, so that $K^\vee K$ is the ring of $\Z_p$--valued functions on $\Z_p^\times$.  Turning to cohomology, it follows by the universal coefficient spectral sequence that $K^0 K = \Hom(\cts(\Z_p^\times, \Z_p), \Z_p)$ and that $K^1 K = 0$.  These correspondences behave as follows:
\begin{enumerate}
    \item The Kronecker pairing \[\S^0 \xrightarrow{c} K \sm K \xrightarrow{1 \sm f} K \sm K \xrightarrow{\mu} K\] is computed by the evaulation pairing \[(c \in K^\vee K, f \in K^0 K) \mapsto f(c).\]
    \item The stable operation $\psi^\lambda$ attached to $[\lambda] \in \Aut \G_m$ is evaluation at $\lambda$.
    \item The stable cooperation $v^{-k} \sm v^k \in \pi_0 K \sm K$ corresponds to the polynomial function $x \mapsto x^k$, as justified by the computation \[\operatorname{ev}_{\lambda}(v^{-k} \sm v^k) = \frac{\psi^\lambda v^k}{v^k} = \frac{\lambda^k v^k}{v^k} = \lambda^k.\]
\end{enumerate}

\noindent These last two facts mean that the behavior of a stable operation on homotopy is identical information to the values of a functional $f$ on the standard polynomial functions $x^k$.  We record this algebraic model as follows:
\begin{lemma}
For any $N \ge 0$, the assignment \[\Hom(\cts(\Z_p^\times, \Z_p), \Z_p) \xrightarrow{(f(x \mapsto x^k))_k} \prod_{k \ge N} \Z_p\] is injective.  A sequence $(x_k)$ is said to be a \emph{K\"ummer sequence} when it lies in this image.\footnote{A bit more explicitly: $(x_k)$ is K\"ummer when for all $h(x) = \sum_{k=N}^n a_k x^k \in \Q[x]$ we have $\sum_{k=N}^m a_k x_k \in \Z_p$.} \qed
\end{lemma}

\begin{remark}
An interesting feature of the Lemma is the auxiliary index $N$, which is \emph{not} part of the property of being K\"ummer.  In $p$--adic geometry, this is reflected by the $p$--adic convergence of the sequence \[d + (p-1)p^r \xrightarrow{r \to \infty} d,\] and hence the continuous reconstruction property \[x_d = \lim_{r \to \infty} x_{d + (p-1)p^r}.\]  In homotopy theory, this is reflected by the reconstruction property $K \sm K[2k, \infty) \simeq K \sm K$.
\end{remark}

\begin{remark}
With this computation in hand, the $p$--local operations $KU_{(p)} \sm KU_{(p)}$ can be recovered from arithmetic fracture, as can the global operations $KU \sm KU$.  The answer is quite similar: $\pi_0 KU \sm KU$ is populated by rational polynomials which evaluate to integers on all integer inputs, called \textit{numerical polynomials}.
\end{remark}

We now pass from $KU$ to $KO$.  To begin, use the Tate trick
\begin{align*}
K \sm KO & \simeq K \sm (K^{hC_2}) & \text{($KO$ is a homotopy fixed point spectrum)} \\
& \simeq K \sm (K_{hC_2}) & \text{(Tate objects vanish $K(1)$--locally)} \\
& \simeq (K \sm K)_{hC_2} & \text{(homotopy colimits pull past smash products)} \\
& \simeq (K \sm K)^{hC_2}, & \text{(Tate objects vanish $K(1)$--locally)}
\end{align*}
so that $\pi_0 K \sm KO = \cts(\Z_p^\times / C_2, \Z_p)$.  Taking fixed points again, we then also have $\pi_* KO \sm KO = \cts(\Z_p^\times / C_2, KO_*)$, and $KO^* KO$ is the $KO_*$--linear dual.  It follows that $[\Susp^{-1} KO, KO] = 0$ and that $[KO, KO] = \Hom(\cts(\Z_p^\times / C_2, \Z_p), \Z_p)$ is torsion-free, which account for our outstanding claims.


\subsubsection{Mazur's construction of Kubota--Leopoldt $p$--adic $L$--functions}

Having learned enough about $KO$--operations to justify the program enacted in the previous subsections, we now need to show that there exist sequences of $p$--adic integers satisfying those criteria.

\begin{theorem}[Mazur]
For any auxiliary $c \in \Z_p^\times$, there is a functional $f_c$ satisfying\footnote{\[\text{Explicitly,\;} f_c(h) = \int_{\Z_p^\times} h(x) d\mu_c = \lim_{r \to \infty} \frac{1}{p^r} \sum_{\substack{0 \le i < p^r \\ p \nmid i}} \int_i^{ci} \frac{h(t)}{t} dt.\]}\footnote{With considerable effort, this output can be halved~\cite[Section 10.3]{AHR}.}\footnote{It also satisfies the normalizing property $\int_{\Z_p^\times} d\mu_c = \frac{1}{p} \log(c^{p-1})$.} \[f_c(x^{k \ge 1}) = \frac{-B_k}{k}(1 - p^{k-1})(1 - c^k).\]
\end{theorem}

\noindent This Theorem is stated in exactly the generality it was originally proven, and so uou might wonder why Mazur had already proven \emph{exactly} what we needed.  To understand his program, recall these two facts about $\zeta$:
\begin{enumerate}
    \item Except for a real Euler factor, $\zeta$ is basically the Mellin transform of the measure $\frac{dx}{e^x - 1}$ (i.e., its sequence of moments): \[\zeta(s) = \frac{1}{\Gamma(s)} \int_0^\infty x^{s-1} \frac{dx}{e^x - 1}.\]
    \item For any $k \in \Z_{> 0}$, $\zeta(1 - k) = -B_k / k$, where $\frac{t}{e^t - 1} = \sum_{k=0}^\infty B_k \frac{t^k}{k!}$.
\end{enumerate}
Mazur's idea was to build a $p$--adic $\zeta$--function by investigating similar $p$--adic integrals, beginning with certain finitary approximations to this one.  To begin, a Bernoulli polynomial for $k \in \Z_{>0}$ is \[\sum_{k=0}^\infty B_k(x) \frac{t^k}{k!} = \frac{t e^{tx}}{e^t - 1}.\]  These polynomials beget Bernoulli distributions according to the rule
\begin{align*}
\Z/p^n\Z & \xrightarrow{E_k} \Q \subseteq \Q_p \\
x \in [0, p^n) & \mapsto k^{-1} p^{n(k-1)} B_k(x p^{-n}).
\end{align*}
A distribution in general is a function on $\Z_p$ such that its value at any node in the $p$--adic tree is equal to the sum of the values of its immediate children, and the $p$--adic integral of a locally constant function with respect to such a distribution is defined by their convolution.  For example, the constant function $1$ factors through $\Z/p$, hence \[\int_{\Z_p} dE_k = \overset{\text{non-obvious}}{\overbrace{\frac{1}{k} \sum_{a=0}^{p-1} B_k\left(\frac{a}{p}\right) = \frac{B_k(0)}{k}}} = \frac{B_k}{k}.\]

However, this distribution is not a \emph{measure}, in the sense that it is not bounded and hence does not extend to a functional on all continuous functions (rather than just locally constant ones).  The standard fix for this is called \emph{regularization}: pick $c \in \Z$ with $p \nmid c$, and set $E_{k,c}(x) = E_k(x) - c^kE_k(c^{-1}x)$.  This is a measure, and for $k \ge 1$ it has total volume given by \[\int_{\Z_p} dE_{k,c} = \int_{\Z_p} dE_k - c^k \int_{\Z_p} dE_k(c^{-1}x) = \frac{B_k}{k}(1 - c^k).\]

These measures interrelate: $E_{k, c} = x^{k-1} E_{1, c}$, and hence the single measure $E_{1, c}$ has all of these values as moments.  We would like to perform $p$--adic interpolation in $k$ to remove the restriction $k \ge 1$, but this is not naively possible: if $k = 0$, say, then we naively have $E_{0, c} = x^{-1} E_{1, c}$, which will not make sense whenever $x \in p\Z_p$.  This is most easily solved by restricting $x$ to lie in $\Z_p^\times$, which has a predictable effect for $k \in \Z_{> 0}$:
\begin{align*}
\int_{\Z_p^\times} x^{k-1} dE_{1,c} & = \int_{\Z_p} x^{k-1} dE_{1,c} - \int_{p\Z_p} x^{k-1} dE_{1, c} \\
& = \int_{\Z_p} x^{k-1} dE_{1,c} - p^{k-1} \int_{\Z_p} x^{k-1} dE_{1, c} \\
& = \frac{B_k}{k}(1 - c^k)(1 - p^{k-1}).
\end{align*}
Hence, the Mellin transform of the measure $dE_{1,c}$ on $\Z_p^\times$ gives a sort of $p$--adic interpolation of the $\zeta$--function.

It also has \emph{exactly} the properties we need to guarantee the existence of an $E_\infty$ orientation $M\Spin \to KO$.  It is remarkable that the three factors in \[\int_{\Z_p^\times} x^{k-1} d E_{1, c} = \frac{B_k}{k} (1 - c^k)(1 - p^{k-1})\] have discernable provenances in the two fields.  In stable homotopy theory these arise respectively in the characteristic series of the orientation $MU \to KU$, in the finite Adams resolution for the $K(1)$--local sphere, and in the Rezk logarithm.  In $p$--adic analytic number theory, they arise as the special values of the $\zeta$--function, the regularization to make it a measure, and the restriction to perform $p$--adic interpolation.  It is completely mysterious how or if these operations correspond.

\begin{remark}
These Bernoulli sequences are \emph{not} the only sequences satisfying these reconstruction properties---in fact, there are infinitely many, and an explicit presentation of them is available~\cite{SprangNaumann}.  Sprang leaves open whether there is a way to single out the Bernoulli solution among the rest, and it seems plausible that this is the only solution with a ``reasonable'' growth rate (as measured in $\R$).  It would also be great if this ``real place'' condition had something to do with a smooth cohomology theory like differential real $K$--theory.
\end{remark}


\subsubsection{Footnotes on the $\tmf$ case}

The case of the orientation $M\String \to \tmf$ has all of the same trappings, but its order of complexity is $(-)^{3/2}$ of the above case, essentially because the height $1$ chromatic fracture \emph{square} gets replaced by the height $2$ chromatic fracture \emph{cube}.  (There is also the issue of the more complicated coefficient ring $\tmf_*$ over $KO_*$.)  Many of the steps remain the same:
\begin{enumerate}
    \item Begin with a rational orientation, which is basically the Witten genus valued in holomorphic expansions of modular forms.
    \item Analyze the homotopy type of $\widehat L_1 \tmf$ and compare it to that of $KO$.  This lets us use another universal coefficient theorem to lift our description of $KO^* KO$ as $KO^*$--valued measures to $\widehat L_1 \tmf^* KO$ as $\widehat L_1 \tmf^*$--valued measures.
    \item The homotopy type of $\widehat L_2 \tmf$ is ``naively irrelevant'' in the chromatic fracture square: maps $b\Spin \to \widehat L_2 \tmf$ factor through $\widehat L_2 b\Spin = \widehat L_2 KO_p = 0$.
    \item However, the logarithm's presence in the chromatic fracture square $\widehat L_1 \tmf \to \widehat L_1 \widehat L_2 \tmf$ has a real effect that must be understood.  This is not easy: the height $2$ logarithm is not so accessible, so this requires a real understanding of power operations in $\tmf$.
    \item You also have to calculate the Miller invariant associated to $\tmf$.  In the case of $E_\infty(M\Spin, KO)$, one uses the $A_\infty$ orientation $MU \to KU$, as well as an understanding of the maps $\pi_* BU \to \pi_* BO$ and $\pi_* KO \to \pi_* KU$.  The case of $E_\infty(M\String, \tmf)$ is similar: one constructs an $A_\infty$ lift of the $\sigma$--orientation $MU[6, \infty) \to K^{\Tate}$, as well as an understanding of the maps $\pi_* BU[6, \infty) \to \pi_* B\String$ and $\pi_* \tmf \to \pi_* K^{\Tate}$.
    \todo{I think the conclusion here is that you have to be a $q$--expansion of a modular form (of a particular weight) with constant term a Bernoulli number and every other coefficient integral.  Mike told me that this (or something like this) fully determines these generalized Eisenstein series; that's nice.}
    \todo{Mike said that there's some kind of Bousfield--Kuhn argument you can make to use the $H_\infty$ orientation to detect the generalized Eisenstein congruences.  I can't figure out what it would be.}
    \item Finally, you have to ramp up the algebraic part of the calculation by identifying the analogues of the Mazur moments in $\pi_* \widehat L_1 \tmf$.  These turn out to be normalized Eisenstein series.
\end{enumerate}

\begin{remark}
The presence of such interesting arithmetic invariants (Bernoulli numbers, Bernoulli polynomials, generalized Eisenstein series, \ldots) hiding in the Miller invariant and its analogues is very striking.  One wonders what the analogous values are (or perhaps the values stemming from the iterated $S^1$--transfer of Baker et al~\cite{BCGHRW}) associated to a Morava $E_\Gamma$.
\end{remark}

\begin{remark}
Some more open questions about this can be found in 
\todo[inline]{Mike has some open questions about the end of this analysis (and in particular about the fiber of the Atkin map that appears in the $K(1)$--local analysis of $\tmf$, an analogue of the chromatic splitting fiber) at the end of his talk notes \textit{The $\String$ orientation of $\tmf$}.  Some of that should be copied here.}
\end{remark}

\todo{$KU$ is known to have a unique $E_\infty$ structure by work of Baker--Richter.  Is this also true of $K^{\Tate}$?  If so, it lends a lot of credibility to this Miller invariant calculation and its relation to $\tmf$.}
\todo{I think it's possible to show, as a side-example, that the total exterior power operation $\lambda^q\co K \to K\ps{q}$ is an $E_\infty$ map where $K\ps{q}$ is $K^{\Tate}$ and \emph{not}, e.g., $K^{\CP^\infty}$.}

















